### 2022-01-07 thesis/chapters/01-introduction.tex

\chapter{Introduction}
\label{chap:introduction}

\begin{quote}
    \textit{The Cephalopod retreats and crawls back up against the glass of the fish tank
        evaluating the situation in front of him. In front of him is a glass jar with
        prey inside. The lid is shut tightly. Never before has he encountered such an
        obstacle in nature, prohibiting him from feeding. He curiously swims closer,
        feeling the shape of the jar with his tentacles. He wraps his tentacles over the
        jar and with a small turn and pop, he releases the lid. The experiment is
        repeated and without hesitation, without failure, he opens the jar, faster, more
    efficiently, every time.}
\end{quote}

\Ac{ML} is one of the most popular fields of research in \ac{AI} studies today.
In recent years, \ac{ML} research has seen some notable achievements in the
academia \cite{ref:lecun:2015, ref:glorot:2010, ref:goodfellow:2014,
ref:quoc:2017}, as well as the industry at large \cite{ref:silver:2016,
ref:silver:2017, ref:zoph:2017, ref:lewis:2017}.  \ac{ML} research has grown
tremendously over the past decade with successes like AlphaGo, which set new
standards for \ac{AI} capabilities by beating the world's best Go player, Lee
Sedol, 4-1 \cite{ref:san-hun:2016}.  Furthermore, an improvement on AlphaGo,
called AlphaZero, learned to play Go, \textit{tabula rasa} \footnote{Latyn word
meaning to learn from no prior knowledge} and managed to beat AlphaGo, 100-1
\cite{ref:silver:2017}.

Historically, \ac{ML} models were built with a specific purpose in mind and
revolved around specialised applications. A notable development was that of Deep
Blue by IBM \cite{ref:keene:1996}, which managed to play chess at a grand master
level. Deep Blue was able to finish 2-4 against the greatest chess player at the
time, Gary Kasparov. Although Deep Blue lost the tournament, this was a major
breakthrough in the field of \ac{AI}. However, Deep Blue was only able to play
chess and could not be utilised for any other purposes \cite{ref:kelley:2010}.
Today, DeepMind's \ac{RL} algorithm is capable of playing all of the original
Atari games \cite{ref:mnih:2013}, but is still very limited to a small set of
problems that it can solve.

Over the past few years, modern hardware capabilities have improved to the
point where workloads in the field of \ac{ML} that were previously computationally
infeasible, is now possible. One such subfield of \ac{ML} is \acp{ANN}. With
the improvement of hardware came an influx of \ac{ML} researchers that focussed
their attention on the training and generalisation capabilities of \acp{ANN}.

A popular field of focus for studing \acp{ANN} is the training process by which
these models are trained. Today, the most common form of training is
\index{supervised training}supervised learning of \acp{FFNN}, a specific type of
\ac{ANN} \cite{ref:reed:1999}. Training of \acp{ANN} is seen as an optimisation problem.
A search algorithm known as a \index{heuristic}\textit{heuristic}
\cite{ref:pearl:1984} is used to assign the optimal values to the \ac{ANN}
weights such that a specified objective function is minimised.

Although the landscape of what can be solved using \acp{ANN} today
is extensive, there still exists no single model that can be generalised to
solving multiple problem classes, across multiple problem domains. This problem
is known as the \ac{NFL} \cite{ref:wolpert:1997}.

Various attempts have been made to improve the generalisation capabilities of
\acp{ANN}. Examples of techniques used include weight \index{dropout}dropout
\cite{ref:srivastava:2014}, \index{weight decay}weight decay
\cite{ref:krogh:1992}, \index{meta-learning}meta-learning of learning parameters
such as the \index{learning rate}learning rate and \index{momentum}momentum of
gradient-based heuristics \cite{ref:zeiler:2012, ref:lv:2017, ref:darken:1992}
and carefully selecting training data. Other techniques focus on improving
generalisation through model design (architecture and topology). Examples of such
techniques include learning of \ac{ANN} architectures through pruning techniques
\cite{ref:cibas:1996, ref:engelbrecht:1996} or constructive techniques
\cite{ref:hassibi:1994, ref:lecun:1990}, adaptive activation functions
\cite{ref:engelbrecht:1995, ref:fletcher:1994}, and quantum methods
\cite{ref:wan:2017, ref:ricks:2004}.

One promising approach to improving \ac{ANN} generalisation capabilities is the
idea of hybridisation \cite{ref:burke:2013}, which makes use of a combination of
multiple learning paradigms in order to yield better generalisation. For
example, heteregenous approaches have been shown to balance the tradeoff between
exploration and exploitation. \cite{ref:nepomuceno:2013}. This hints towards the
idea that \ac{ML} models could yield better generalisation capabilities if the
learning process and learning mechanism applied is not statically defined, but
rather dynamic and under the control of some control mechanism.

The learning process is influenced by various factors, including the type of
learning technique used, how the model parameters are initialised, the
hyper-parameters used, the constraints of learning and the learning mechanism
itself. In order to yield better generalisation capabilities, the model should
be trained by a learning technique that has at its disposal a more diverse set
of learning behaviours from which it can dynamically select should the need arise
\cite{ref:huang:2009}. In doing so, balancing between exploration and
exploitation of solutions is a dynamic process that would yield better
generalisation capabilities.

An example of dynamically balancing exploitation and exploration during the
search process is done by dynamically adjusting and learning the heuristic
\index{hyper-parameters}\textit{hyper-parameters} as part of the learning
process. This field of study is known as \index{meta-learning}meta-learning
\cite{ref:giraud:2004}. \index{meta-learning}Meta-learning of heuristic
hyper-parameters as applied in the training of \acp{ANN} has shown to yield good
generalisation results \cite{ref:hospedales:2020, ref:vilalta:2002}.

A more recent suggestion related to the field of
\index{meta-learning}meta-learning is to dynamically select and/or adjust the
heuristic used throughout the training process. This approach is about the
hybridisation of heuristics as they are applied to some optimisation problem.
These methods are referred to as \acp{HH} and focus on finding the best
heuristic in \textit{heuristic-space}. One such form of \ac{HH} is a
population-based approach that guides the search process by automatically
selecting heuristics from a heuristic-space to be applied to a collection of
different candidate solutions in the solution-space. This collection of
candidate solutions are referred to as a \textit{population} of
\textit{entities}, where each \textit{entity} is a single candidate solution to
the problem being optimised.

Population-based \acp{HH} implement the strategy of multiple heuristics working
together to solve a problem. However, finding the best heuristic to use is
non-trivial and some \textit{selection strategy} must be used to select the best
heuristic. \acp{HH} that implement such a selection strategy is referred to as
\textit{selection} \acp{HH}. The term \textit{selection} refers to the ability
of the \ac{HH} to select the best heuristic from a pool of low-level heuristics.

These selection \acp{HH} can be further categorised based on what information
they use in the selection strategy. One specific type of selection \ac{HH} is
called a \index{multi-method populated-based meta-heuristic}\textit{multi-method
population-based \index{meta-heuristic}meta-heuristic}
\cite{ref:vanderstockt:2018}. The term \index{multi-method}\textit{multi-method}
refers to the incorporation of different low-level heuristics, with different
search behaviours, into the heuristic-space. The term \textit{population-based}
refers to the utilisation of a population of entities that represent candidate
solutions. Finally, the term \index{meta-heuristic}\textit{meta-heuristics}
refers to the \ac{HH} as a heuristic that does not have any domain knowledge and
only makes use of information from the search process.
\citeauthor{ref:grobler:2015} \cite{ref:grobler:2015} mentioned that \acp{HH}
have been shown to solve a number of problems including bin-packing, examination
timetabling, production scheduling, the traveling salesman problem, vehicle
routing problem and many more.

This thesis focusses on the development of a \ac{HH} to train \acp{FFNN} in a
supervised learning approach.

\section{Problem Statement}
\label{sec:introduction:problem}

Many different heuristics have been developed and used to train \acp{FFNN}
\cite{ref:gudise:2003, ref:rakitianskaia:2012, ref:montana:1989}.  The problem
lies in the process of finding the best heuristic to train the \ac{FFNN}. This
process is often non-trivial and could be a time-consuming exercise.  Consider
that selection of the best heuristic as applied to optimisation problems, such
as training \acp{FFNN}, is very problem dependent \cite{ref:allen:1996,
ref:drake:2020, ref:pillay:2018}. Furthermore, the \ac{NFL} shows that no single
heuristic performs better than any other heuristic for all problems
\cite{ref:wolpert:1997}.

Careful, systematic selection is thus required to find and select the best
heuristic to train \acp{FFNN}. In the past, researchers selected the best
heuristic by trail and error. A set of heuristics and carefully selected
hyper-parmeters would be implemented, followed by an empirical test to evaluate
the performance of each heuristic for a given problem domain
\cite{ref:pillay:2015}. In this way, researchers were able to determine which
heuristics and which hyper-parameters performed well for different problems.
However, this approach is problematic as it is time-consuming and tedious.

\section{Motivation}
\label{sec:introduction:motivation}

A process is required to alleviate the burden of having to exhaustively test
each implementation of heuristic and hyper-parameters, for every problem being
optimised.

A modern approach is to use \acp{HH} to automated the process for selecting the
best heuristic when applied to some optimisation problem. The best heuristic
might not be a single heuristic, but rather a hybridisation of heuristics
\cite{ref:pillay:2015}. Therefore, careful consideration is required to include
a diverse and applicable set of heuristics to select from that is capable of
solving the given problem. \citeauthor{ref:pillay:2015} \cite{ref:pillay:2015}
mentions that only heuristics that are known to have the potential to solve the
problem must be included in the set of heuristics to select from.

\acp{HH} have been implemented on a wide range of optimisation problems
\cite{ref:grobler:2011, ref:burke:2013, ref:pillay:2015}, but no research has
been found where \acp{HH} have been applied on the training of \acp{FFNN}.

The training of \acp{FFNN} is seen as a computational search problem. \Acl{HH}
as defined by Burke et al. \cite{ref:burke:2010} is ``a search method or
learning mechanism for selecting or generating heuristics to solve computational
search problems". \acp{HH} is a field of research aimed at the automated
selection, combination, adaptation and generation of multiple lower-level
heuristics to efficiently solve computational search problems. This implies that
one should be able to apply \acp{HH} in the context of training \acp{FFNN}.

\citeauthor{ref:grobler:2012} \cite{ref:grobler:2012} mentions that the focus of
\acp{HH} is on automating the development of the \textit{learning mechanism}
used to obtain an appropriate solution for an optimisation problem. The authors
go on to explain that \acp{HH} employ a high-level heuristic that focusses on
finding the best low-level heuristic in heuristic-space that could include
\textit{single-method} and \textit{multi-method} optimistion algorithms
(heuristics). Applying \acp{HH} to train \acp{FFNN} could generally be applied
to multiple problems given that the set of low-level heuristics include
heuristics that have the potential to solve the problem at hand
\cite{ref:burke:2010}. This relieves the burden of having to select the best
heuristic to use by trial and error and is generally applicable to multiple
problems. The automation of the selection process as well as the generalisation
capabilities of \acp{HH}, if applied in the context of training \acp{FFNN},
could reduce the effort required to find the best heuristic to use.

Note that training \acp{FFNN} using \acp{HH} is not to be confused by the
training of \acp{FFNN} using \index{ensemble networks}\textit{ensemble networks}
or \index{query by committees}\textit{query by committees}.
\citeauthor{ref:pappa:2014} \cite{ref:pappa:2014} describes \index{ensemble
networks}ensemble networks as a combination method in
\index{meta-learning}meta-learning whereby multiple \acp{ANN} are jointly used
to solve a problem. Each member network in the ensemble gets trained using a
specified heuristic and generally this heuristic is applied to the same member
throughout the training process. The results of all of the member networks are
then joined together using some mechanism \cite{ref:zhou:2002}. This mechanism
could be weighted averages or majority or weighted voting. \index{ensemble
networks}Ensemble networks do not search through the heuristic space and do not
reselect lower-level heuristics for each member, while \acp{HH} do. There is
thus no hybridisation of heuristics in \index{ensemble networks}ensemble
networks and therefore, no information-sharing between any of the members of the
ensemble, in comparison to \acp{HH} that do incorporate some shared information
space.

Many different types of \acp{HH} have been implemented and shown to solve many
different problems. Some notable examples include the
\index{tabu-search}tabu-search \ac{HH} of Burke et al. \cite{ref:burke:2010} and
the \index{simulated annealing}simulated annealing-based \ac{HH} by Dowsland et
al. \cite{ref:dowsland:2007}.

This research takes a particular interest in developing a high-level heuristic
that makes use of probability theory and Bayesian statistical concepts to guide
the heuristic selection process. This research develops the novel \Ac{BHH}, a
new high-level heuristic that utilises \index{Bayesian analysis}\textit{Bayesian
analysis} as the selection mechanism for a \ac{HH} to select the appropriate
heuristic to train \acp{FFNN}.

\section{Objectives}
\label{sec:introduction:objectives}

The main objectives of this research is focussed on developing a novel \Ac{BHH}
selection mechanism in a \ac{HH} framework that can be used to train \acp{FFNN}.
In order to reach this objective, the following sub-objectives are defined:

\begin{itemize}
    \item
    Conduct a literature study on \acp{ANN} in order to provide
    the necessary background information of the \ac{AN}, \acp{FFNN} and the
    training process.

    \item
    Conduct a literature study on different types of heuristics that have been
    used to train \acp{FFNN}. The literature study will provide the necessary
    background information to understand how the these different heuristics can
    be used in the set of heuristics to be selected to train the \ac{FFNN}.

    \item
    Conduct a literature study on \index{meta-learning}meta-learning and
    \acp{HH} in order to provide the required background information
    necessary to propose and develop a new high-level heuristic.

    \item
    Conduct a literature study on probability theory and Bayesian statistics
    such as Bayesian inference and Bayesian analysis to provide the necessary
    background information required to understand how Bayesian approaches can be
    used as a learning technique.

    \item
    Develop a novel \Ac{BHH} selection mechanism for a \ac{HH} that makes use
    of Bayesian statistics to guide the \ac{HH} search process while training
    \acp{FFNN} on different problems.

    \item
    Conduct an empirical study to show that the developed \Ac{BHH} can
    effectively be used to train \acp{FFNN}.

    \item
    Conduct an empirical study and critically evaluate the performance of the
    developed \Ac{BHH} compared to individual heuristics in the heuristic-space
    as they are used to train \acp{FFNN} on different problems.

    \item
    Conduct an empirical study and critically evaluate the performance of the
    developed \Ac{BHH} as a selection mechanism to other selection mechanisms
    for \acp{HH} as they are used to train \acp{FFNN} on different problems.

    \item
    Conduct an empirical study that investigate variations of the \Ac{BHH} and the
    effects of design decisions and hyper-parameters on the search process.

    \item
    Provide a statistically sound analysis of the results obtained from the
    empirical study.
\end{itemize}

\section{Contributions}
\label{sec:introduction:contributions}

The results obtained from this research contributes to the field of study in the
following ways:

\begin{itemize}
    \item
    To the knowledge of the author, this study provides the first application
    of using \acp{HH} to train \acp{FFNN}.

    \item
    A novel heuristic selection operator is used that focusses on using Bayesian
    statistics to calculate the probability that a heuristic should be selected
    in order to efficiently train \acp{FFNN}. The resulting \ac{HH} is referred
    to as the \Acl{BHH}.

    \item
    The results of the empirical study show that the \Ac{BHH} perform generally
    better than the individual lower-level heuristics as applied to the same
    problems, with statistical significance and certainty.

    \item
    The results of the empirical study show that the \Ac{BHH} is able to select
    the best heuristic to train \ac{FFNN} in general, with statistical
    certainty. This relieves researchers from the burden of having to do this
    selection process manually through trial and error.

    \item
    The results of the empirical study show that the \Ac{BHH}, given a diverse
    set of lower-level heuristics, will generally produce good results when
    applied to multiple problems.

    \item
    Finally, the results of the empirical study show that the \Ac{BHH} is
    capable of utilising \textit{a priori} \footnote{Latin word, meaning "from
    what comes before"} knowledge in which a predefined selection bias is used
    for heuristics the are known to be well suited for certain problems.
\end{itemize}

\section{Thesis Outline}
 \label{sec:introduction:outline}

The remainder of this thesis is structured as follows:

 \begin{itemize}
 	\item
    \textbf{Chapter~\ref{chap:anns}} provides a literature study on \acp{ANN}
    and the various components that make up an \ac{AN}. The focus is on the
    training of \acp{FFNN} and it is shown how the training of \acp{FFNN} is
    seen as an optimisation problem.

	\item
    \textbf{Chapter~\ref{chap:heuristics}} provides details on various types of
    heuristics and metaheuristics that have been used to train \acp{FFNN}. A
    literature study is done to provide details on the search behaviours, application and
    implementation of each heuristics in the context of training \acp{FFNN}.

	\item
    \textbf{Chapter~\ref{chap:hhs}} presents a literature study on the details
    of \acp{HH} and \index{meta-learning}meta-learning in general. A
    discussion follows on the current landscape of \ac{HH} research. It is shown
    how \acp{HH} is a suitabe approach for training \acp{FFNN}.

    \item
    \textbf{Chapter~\ref{chap:probability}} presents a literature study on
    probability theory. Probability distributions and
    conjugate priors are discussed in detail. The chapter concludes with a
    detailed discussion on Bayesian statistics, specifically focussing on
    Bayesian inference and analysis.

	\item
    \textbf{Chapter~\ref{chap:bhh}} presents the developed \Ac{BHH}. It is shown
    how the \Ac{BHH} is implemented as a selection mechanism in the context of a
    \ac{HH} framework. The \Ac{BHH} is shown to implement a \index{Na\"ive Bayes
    classifier}Na\"ive Bayes classifier. The probabilistic model that is
    implemented is derived and discussed in detail. Finally, the chapter
    concludes with a detailed discussion on how \index{Bayesian
    analysis}Bayesian analysis is used to guide the heuristic selection process.
    The update step (training step) for prior probability concentration
    parameters is derived and discussed in detail. The \Ac{BHH} algorithmic
    implementation, variants and suggested application are presented as well.

	\item
    \textbf{Chapter~\ref{chap:methodology}} presents a detailed description of
    the empirical process and the setup of each experiment. It discusses the
    datasets used, the \ac{FFNN} architecture and topology used, heuristics that
    are used, the configuration of hyper-parameters, initialisation techniques
    used, performance measures used and other smaller details to the
    implementation of each experiment.  Finally discussions follow on how the
    results are analysed. This includes a discussion on the method used to
    determine statistical significance.

    \item
    \textbf{Chapter~\ref{chap:results}} provides and discusses the results of
    the empirical study, in detail. A baseline comparison is done by comparing
    the performance of the \Ac{BHH} to that of all the individual lower-level
    heuristics on all the datasets. These datasets include classification and
    regression problems of various sizes and complexity.  Detailed results are
    presented on the performance of the \Ac{BHH} as a selection mechanism for
    \acp{HH} and a brief comparison is made to other selection mechanisms.
    Finally, results are presented on the effects of the hyper-parameters of the
    \Ac{BHH} on the training process. Discussions follow on how the \Ac{BHH} is
    shown to automate the selection of the best heuristic during the training of
    \acp{FFNN}, as applied to a range of problems, alleviating the burden on
    researchers to apply traditional trial and error approaches. The chapter is
    concluded with a brief argument in favor of the general applicability of the
    \Ac{BHH} to more problem domains.

	\item
    \textbf{Chapter~\ref{chap:conclusion}} summarises the research done in this
    thesis along with a brief overview of findings made throughout the research
    process. A review of the research goals are given and suggestions for future
    research are made.
\end{itemize}

This thesis is accompanied by a full index given at page~\pageref{index} along with the following appendices:

\begin{itemize}
	\item
    \textbf{Appendix~\ref{app:datasets}} provides details on the datasets used
    for the empirical analysis.

 	\item
    \textbf{Appendix~\ref{app:acronyms}} provides a list of the important
    acronyms used or newly defined in the course of this work, as well as their
    associated definitions.

 	\item
    \textbf{Appendix~\ref{app:symbols}} lists and defines the mathematical
    symbols used in this work, categorised according to the relevant chapter
    in which they appear.

 	\item
    \textbf{Appendix~\ref{app:derived_publications}} lists the publications
    derived from this work.
 \end{itemize}

To best view the illustrations, tables and figures presented throughout this
thesis, it is recommended that the thesis be viewed in colour.