### 2022-01-07 thesis/latex/packages.tex

\usepackage{acro}                   % American Mathematical Society environments}
\usepackage{amsmath}                % American Mathematical Society environments}
\usepackage{amssymb}                % American Mathematical Society Symbols
\usepackage{amsthm}                 % American Mathematical Society Theorems
% \usepackage{algorithm}            % Typeset algorithm captions
\usepackage{booktabs}
\usepackage[ruled,vlined,linesnumbered,noresetcount]{algorithm2e}
\usepackage{bbm}                    % Blackboard variants of Computer Modern Fonts
\usepackage[sorting=none]{biblatex} % BibLaTeX referencing
\usepackage{caption}                % Improved customisation of captions
\usepackage{colortbl}               % Adds colors to tables
\usepackage{datetime}               % Various date operations
\usepackage{fancyhdr}               % Add fancy headers and footers
\usepackage{float}                  % Improved interface for adding floating objects
\usepackage[T1]{fontenc}            % Allows user to select font encodings
\usepackage[bottom]{footmisc}
\usepackage[
a4paper,                            % Paper size
width=150mm,                        % Paper width
top=25mm,                           % Top margin
bottom=25mm,                        % Bottom margin
bindingoffset=6mm,                  % Binding margin
showframe                           % Adds frame to see hbox and vbox
]{geometry}                         % Easy and flexible user interface to customize page layout,
\usepackage[pdftex]{graphicx}                   
\usepackage{ifpdf}                  % Selective setup for PDF compilation
\usepackage[utf8]{inputenc}         % Set the input encoding to UTF-8
\usepackage{imakeidx}               % A general purpose hierarchical index generator
\usepackage{listings}               % Typeset sourcecode listings
\usepackage{lmodern}
\usepackage[final]{microtype}
\usepackage{pgfplots}               % Plot functions
\usepackage{relsize}
\usepackage{subcaption}             % Adds support for sub-figures and sub-captions
\usepackage{texnames}               % BIBTeX, SliTeX, AMSTeX, PiCTeX and TeXsis logos
\usepackage{todonotes}              % Add todo notes functionality
\usepackage{url}                    % URL typesetting
\usepackage{verbatim}               % Multi-line Comments

% Dependency reliant
\usepackage[
bookmarks=true,                         % (option) Build bookmarks
bookmarksnumbered=true                  % (option) Add section numbers to bookmarks
]{hyperref}
\usepackage[all]{hypcap}                % Correct PDF figure links to top of image
\usepackage[ocgcolorlinks]{ocgx2}       % Allows you to create OCGs (Optional Content Groups)