### 2022-01-07 thesis/preamble/quote.tex

\pagestyle{empty}
\vspace*{\fill}
\begin{quotation}
    \noindent ``Men go abroad to wonder at the heights of mountains, at the huge
    waves of the sea, at the long courses of the rivers, at the vast compass of
    the ocean, at the circular motions of the stars and they pass by themselves
    without wondering.''
\end{quotation}
\begin{flushright}
    St. Augustine
\end{flushright}
\vspace{1cm}

\begin{quotation}
\noindent
    ``The measure of intelligence is the ability to change.''
\end{quotation}
\begin{flushright}
    Albert Einstein
\end{flushright}
\vspace*{\fill}
\newpage