### 2022-01-07 thesis/preamble/supervisors.tex

\pagestyle{empty}
\vspace*{\fill}
\noindent
\parbox{4cm}{\textbf{Supervisor:}} Dr. A. Bosman \\*[2mm]
\parbox{4cm}{\textbf{~}} University of the Pretoria \\*[2mm]
\parbox{4cm}{\textbf{~}} Department of Computer Science\\*[1cm]
\parbox{4cm}{\textbf{Co-Supervisor:}} Dr. C.W. Cleghorn \\*[2mm]
\parbox{4cm}{\textbf{~}} University of the Witwatersrand\\*[2mm]
\parbox{4cm}{\textbf{~}} Department of Computer Science\\*[1cm]
\parbox{4cm}{\textbf{Co-Supervisor:}} Prof. A.P. Engelbrecht \\*[2mm]
\parbox{4cm}{\textbf{~}} University of Stellenbosch\\*[2mm]
\parbox{4cm}{\textbf{~}} Department of Industrial Engineering\\*[1cm]
\parbox{4cm}{\textbf{Degree:}} Master of Science, Computer Science (Artificial
Intelligence)
\newpage