### 2022-09-12 dissertation/preamble/acknowledgements.tex

\pagestyle{empty}
\begin{center}
    \Large
    \textbf{Acknowledgements}
\end{center}

\begin{quotation}
    \noindent
    ``I've made the most important discovery of my life. It's only in
    the mysterious equation of love that any logical reasons can be found. I'm
    only here tonight because of you. You are the only reason I am \ldots you are
    all my reasons.''
\end{quotation}
\begin{flushright}
    Prof. J.F. Nash, Jr.
\end{flushright}

\vspace{0.3cm}
\noindent
I would like to thank the following people, without whom this work would never
have been possible:

\begin{itemize}
    \item To my wife, Taylah. You are all my reasons.

    \item To my daughters, Bean\'e and  Aleah. Thank you for giving life meaning.

    \item To my parents, Adr\'e and Engela. Thank you for always believing in me.

    \item To my sister, Jeannie and her husband, Jaco. Thank you for always motivating me.

    \item To my supervisors, Dr. A. Bosman, Prof. C.W. Cleghorn and Prof. A.P. Engelbrecht. Thank you for not giving up on me.

    \item To my creator, God. Thank You \ldots for everything!
\end{itemize}

\vspace{\fill}

\begin{center}
    This project was made possible by the National Research Foundation (NRF) and the Center for High Performance Computing (CHPC) Cluster at the Council for Scientific and Industrial Research (CSIR).
\end{center}

\newpage