### 2022-09-12 dissertation/preamble/publication.tex

\pagestyle{empty}
\vspace*{\fill}
\small

% Publication reference
\noindent
\textsf{Publication:}\\*[2.5mm]
\textsf{
    \fontsize{9}{10pt}
    \selectfont
    \textit{
        Schreuder, A.N. Training Feedforward Neural Networks with Bayesian
        Hyper-Heuristics. Master's Dissertation. University of Pretoria,
        Department of Computer Science, Pretoria, South Africa.
        2022
    }
}
\vspace{1cm}

% Online assets
\noindent
\textsf{Electronic, hyperlinked versions of this dissertation, including
data and source code are available online at:}\\*[2.5mm]
\textsf{
    \fontsize{9}{10pt}
    \selectfont
    \url{https://github.com/arneschreuder/masters.ai}
}
\newpage