### 2022-09-12 dissertation/preamble/titlepage.tex

\pagestyle{empty}
\pagenumbering{Alph}
\begin{titlepage}
    \begin{center}
        % Logo
        \includegraphics[width=0.5\textwidth]{up_logo.eps}
        \vspace{2cm}

        % Title
        \Huge
        \textsf{Training Feedforward Neural Networks with Bayesian
        Hyper-Heuristics}
        \vspace{1cm}

        \large
        \textsf{by}
        \vspace{1cm}

        % Author
        \large
        \textsf{A.N. Schreuder}
        \vfill

        % Degree
        \normalsize
        \textsf{
            Submitted in partial fulfilment of the requirements for the degree\\
            Master of Science, Computer Science (Artificial Intelligence)\\
            in the Faculty of Engineering, Built Environment and Information
            Technology (EBIT)\\
            University of Pretoria,\\
            Pretoria.
        }
        \vspace{2cm}

        % Date
        \normalsize
        \textsf{\monthyeardate\today}
    \end{center}
\end{titlepage}
\newpage