### 2022-09-21 dissertation/preamble/supervisors.tex
✨ (feat): Review, analysis script updated

\pagestyle{empty}
\vspace*{\fill}
\noindent
\parbox{4cm}{\textbf{Supervisor:}} Dr. A. Bosman \\*[2mm]
\parbox{4cm}{\textbf{~}} University of the Pretoria \\*[2mm]
\parbox{4cm}{\textbf{~}} Department of Computer Science\\*[1cm]
\parbox{4cm}{\textbf{Supervisor:}} Prof. C.W. Cleghorn \\*[2mm]
\parbox{4cm}{\textbf{~}} University of the Witwatersrand\\*[2mm]
\parbox{4cm}{\textbf{~}} Department of Computer Science\\*[1cm]
\parbox{4cm}{\textbf{Supervisor:}} Prof. A.P. Engelbrecht \\*[2mm]
\parbox{4cm}{\textbf{~}} Stellenbosch University\\*[2mm]
\parbox{4cm}{\textbf{~}} Department of Industrial Engineering, Computer Science Division\\*[1cm]
\parbox{4cm}{\textbf{Degree:}} Master of Science, Computer Science (Artificial Intelligence)
\newpage