### 2022-11-24 dissertation/preamble/publication.tex

\pagestyle{empty}
\vspace*{\fill}
\small

% Publication reference
\noindent
\textsf{Publication:}\\*[2.5mm]
\textsf{
    \fontsize{9}{10pt}
    \selectfont
    \fullcite{ref:schreuder:2022}
}
\vspace{1cm}

% Online assets
\noindent
\textsf{Electronic, hyperlinked versions of this dissertation, including
    data and source code are available online at:}\\*[2.5mm]
\textsf{
    \fontsize{9}{10pt}
    \selectfont
    \url{https://github.com/arneschreuder/masters.ai}
}
\newpage