### 2022-11-30 dissertation/appendix/datasets.tex
💚 Dedicated to Taylah, Beané and Aleah

\chapter{Datasets}
\label{app:datasets}

This appendix provides insight into the preparation and usage of datasets that were used as part of the empirical process. The section is structured as follows:

\begin{itemize}
	\item \textbf{Section \ref{app:datasets:overview}} provides an overview of the data processing that was executed on each dataset.

	\item \textbf{Section \ref{app:datasets:example}} provides an example data processing script for the iris dataset. The same approach was followed for all datasets.
\end{itemize}

\section{Overview}\label{app:datasets:overview}

This section provides a brief overview of all the data processing operations that were executed on all datasets. The data processing tasks included:

\begin{itemize}
	\item Exploratory analysis and visualisation off data distribution.

	\item Dropping of rows that contain missing data in and standardisation of data types.

	\item Normalising all numerical features using the \index{standard score scaler}standard score scaler and one-hot encoding of all categorical features.

	\item Splitting the datasets into a training set (80\%) and a testing set (20\%) and then shuffling and batching of data.
\end{itemize}

\section{Example}\label{app:datasets:example}

This section provides an example data processing script that was used for the iris dataset. All the datasets used in this dissertation follow the same data processing principles.

\includepdf[pages=-]{attachments/iris.pdf}