### 2022-12-06 dissertation/appendix/symbols.tex
✏️ (final-review): References, datasets, start of symbols.

\chapter{Symbols}
\label{app:symbols}

This appendix summarises the important symbols used throughout the dissertation. For each symbol, a short description is provided as well as the reference to the equation that makes use it. Symbols are divided according to the chapter in which they were introduced and are ordered alphabetically.

\section{Chapter~\ref{chap:introduction}: Introduction}
\label{sec:symbols:introduction}

\section{Chapter~\ref{chap:anns}: Artificial Neural Networks}
\label{sec:symbols:anns}

\begin{description}
	\item
	      [\parbox{1cm}{$\alpha$}] Control parameter for leaky ReLU that controls non-zero gradients. (See Eq.~\eqref{eq:an_function_mapping} p.~\pageref{eq:an_function_mapping})
\end{description}

\section{Chapter~\ref{chap:heuristics}: Heuristics}
\label{sec:symbols:heuristics}

\section{Chapter~\ref{chap:hhs}: Hyper-Heuristics}
\label{sec:symbols:hhs}

\section{Chapter~\ref{chap:probability}: Probability}
\label{sec:symbols:probability}

\section{Chapter~\ref{chap:bhh}: Bayesian Hyper-Heuristic}
\label{sec:symbols:bhh}

\section{Chapter~\ref{chap:methodology}: Methodology}
\label{sec:symbols:methodology}

\section{Chapter~\ref{chap:results}: Results}
\label{sec:symbols:results}

\section{Chapter~\ref{chap:conclusion}: Conclusion}
\label{sec:symbols:conclusion}