### 2022-12-09 dissertation/appendix/symbols.tex
✨ (symbols): Chapter 2 done

\chapter{Symbols}
\label{app:symbols}

This appendix summarises the important symbols used throughout the dissertation. For each symbol, a short description is provided as well as the reference to the equation that makes use it. Symbols are divided according to the chapter in which they were introduced and are ordered alphabetically.

\section{Chapter~\ref{chap:anns}: Artificial Neural Networks}
\label{sec:symbols:anns}

\begin{description}
	% \item
	%       [\parbox{1cm}{$EXAMPLE$}] THIS IS AN EXAMPLE DESCRIPTION. \lbrack Eq.~\eqref{eq:an_function_mapping} pg.~\pageref{eq:an_function_mapping} \rbrack

	\item [\parbox{1cm}{$I$}] The input pattern's dimension size.
	\item [\parbox{1cm}{$T$}] The target pattern's dimension size.
	\item [\parbox{1cm}{$\mathbb{R}$}] The real-number space.
	\item [\parbox{1cm}{$\mathbb{R}^{I}$}] The real-number space in I dimensions.
	\item [\parbox{1cm}{$\mathbb{R}^{T}$}] The real-number space in T dimensions.
	\item [\parbox{1cm}{$f_{AN}$}] The mapping function realised by the \acs{AN}.
	\item [\parbox{1cm}{$\boldsymbol{x}$}] The input pattern/vector.
	\item [\parbox{1cm}{$x_{i}$}] The $i$-th dimension of the input pattern/vector.
	\item [\parbox{1cm}{$\theta$}] The bias unit.
	\item [\parbox{1cm}{$\boldsymbol{v}$}] The weight vector.
	\item [\parbox{1cm}{$v_{i}$}] The $i$-th dimension of the weight vector.
	\item [\parbox{1cm}{$net$}] The net input signal to the \acs{AN}.
	\item [\parbox{1cm}{$f(net)$}] The activation function over net input signal.
	\item [\parbox{1cm}{$\boldsymbol{y}$}] The output vector.
	\item [\parbox{1cm}{$\boldsymbol{y}_{p}$}] The output vector for pattern $p$.
	\item [\parbox{1cm}{$k$}] Index for the output vector.
	\item [\parbox{1cm}{$x_{i,p}^{'}$}] The normalised form of the input at dimension $i$, pattern $p$.
	\item [\parbox{1cm}{$x_{i,p}$}] The original input at dimension $i$, pattern $p$.
	\item [\parbox{1cm}{$x_{i_{min}}$}] The input's minimum value at dimension $i$.
	\item [\parbox{1cm}{$x_{i_{max}}$}] The input's maximum value at dimension $i$.
	\item [\parbox{1cm}{$p$}] Index for the input data patterns.
	\item [\parbox{1cm}{$P$}] Total number of input data patterns.
	\item [\parbox{1cm}{$\boldsymbol{x}_p$}] The $p$-th input vector/pattern.
	\item [\parbox{1cm}{$\mu_i$}] The mean of the input at dimension $i$.
	\item [\parbox{1cm}{$\sigma^2_i$}] The unit variance of the input at dimension $i$.
	\item [\parbox{1cm}{$i$}] Index for the input vector.
	\item [\parbox{1cm}{$\omega$}] The general weight vector used for normalisation.
	\item [\parbox{1cm}{$\omega_{i}$}] The general weight vector used for normalisation at dimension $i$.
	\item [\parbox{1cm}{$\omega_{min}$}] The lower bound of the uniform distribution used for weights initialisation.
	\item [\parbox{1cm}{$\omega_{max}$}] The upper bound of the uniform distribution used for weights initialisation.
	\item [\parbox{1cm}{\footnotesize{$fanin$}}] The number of input neurons to the weight vector.
	\item [\parbox{1cm}{\footnotesize{$fanout$}}] The number of output neurons from the weight vector.
	\item [\parbox{1cm}{$\sigma$}] The standard deviation of the truncated normal distribution used for weight initialisation.
	\item [\parbox{1cm}{$K$}] Number of output units.
	\item [\parbox{1cm}{$x_{i+1}$}] The $i+1$-th dimension of the input pattern/vector.
	\item [\parbox{1cm}{$v_{i+1}$}] The $i+1$-th dimension of the weight vector.
	\item [\parbox{1cm}{$net^{'}$}] The augmented net input signal that includes the bias term.
	\item [\parbox{1cm}{$\tau$}] Phase shift threshold parameter.
	\item [\parbox{1cm}{$f(x)$}] The function used to describe the \acs{LReLU} activation function.
	\item [\parbox{1cm}{$f$}] A shorted form of the activation function.
	\item [\parbox{1cm}{$x^{+}$}] The positive part of the input parameter's domain.
	\item [\parbox{1cm}{$\alpha$}] Scaling parameter used by the \acs{LReLU} activation function.
	\item [\parbox{1cm}{$\lambda$}] Scaling/control parameter used by the sigmoid and tanh activation functions.
	\item [\parbox{1cm}{$x_{min}$}] The input's minimum value used by the min-max scaler.
	\item [\parbox{1cm}{$x_{max}$}] The input's maximum value used by the min-max scaler.
	\item [\parbox{1cm}{$\mathbb{R}^{K}$}] The real-number space in $K$ dimensions.
	\item [\parbox{1cm}{$y_{k}$}] The $k$-th dimension of the output vector.
	\item [\parbox{1cm}{$y^{'}_k$}] The softmax value of the output at dimension $k$.
	\item [\parbox{1cm}{$\boldsymbol{y^{'}}$}] The softmax form of the output vector.
	\item [\parbox{1cm}{$x_{i}$}] The $i$-th dimension of the input vector.
	\item [\parbox{1cm}{$h_{j}$}] The $j$-th dimension of the hidden layer.
	\item [\parbox{1cm}{$j$}] Index used for the hidden layer.
	\item [\parbox{1cm}{$v_{i,j}$}] The weight associated with input node $x_{i}$ and the hidden node $h_{j}$.
	\item [\parbox{1cm}{$w_{j,k}$}] The weight associated with hidden node $h_{j}$ and the output node $y_{k}$.
	\item [\parbox{1cm}{$net_{i,h}$}] The net input signal at the hidden layer.
	\item [\parbox{1cm}{$net_{h,y}$}] The net input signal at the output layer.
	\item [\parbox{1cm}{$\epsilon$}] The cost/loss produced by the loss function.
	\item [\parbox{1cm}{$y_{k,p}$}] The output produced by the \acs{ANN} at the $k$-th dimension for the $p$-th pattern.
	\item [\parbox{1cm}{$\hat{y}_{k,p}$}] The target output for the \acs{ANN} at the $k$-th dimension for the $p$-th pattern.
	\item [\parbox{1cm}{$\beta$}] Mini-batch size.
	\item [\parbox{1cm}{$\mathbbm{1}$}] The indicator function.
	\item [\parbox{1cm}{$c$}] Index used for $c$-th class.
	\item [\parbox{1cm}{$C$}] Total number of classes.

\end{description}

\section{Chapter~\ref{chap:heuristics}: Heuristics}
\label{sec:symbols:heuristics}

\begin{description}
	\item [\parbox{1cm}{$\alpha$}]
	\item [\parbox{1cm}{$\beta_{1}$}]
	\item [\parbox{1cm}{$\beta_{2}$}]
	\item [\parbox{1cm}{$\beta^{t}_{1}$}]
	\item [\parbox{1cm}{$\beta^{t}_{2}$}]
	\item [\parbox{1cm}{$\beta$}]
	\item [\parbox{1cm}{$\boldsymbol{\hat{y}}_{i}(t)$}]
	\item [\parbox{1cm}{$\boldsymbol{G_{t}}$}]
	\item [\parbox{1cm}{$\boldsymbol{g}_{t}^{2}$}]
	\item [\parbox{1cm}{$\boldsymbol{g}_{t}$}]
	\item [\parbox{1cm}{$\boldsymbol{g}^{2}_{t}$}]
	\item [\parbox{1cm}{$\boldsymbol{g}$}]
	\item [\parbox{1cm}{$\boldsymbol{m}_{t}$}]
	\item [\parbox{1cm}{$\boldsymbol{m}_{t+1}$}]
	\item [\parbox{1cm}{$\boldsymbol{r}_{1} \sim U(0,1)^J$}]
	\item [\parbox{1cm}{$\boldsymbol{r}_{2} \sim U(0,1)^J$}]
	\item [\parbox{1cm}{$\boldsymbol{u}_{i}(t)$}]
	\item [\parbox{1cm}{$\boldsymbol{v}_{i}(t)$}]
	\item [\parbox{1cm}{$\boldsymbol{V}_{max}$}]
	\item [\parbox{1cm}{$\boldsymbol{v}_{t}$}]
	\item [\parbox{1cm}{$\boldsymbol{v}_{t+1}$}]
	\item [\parbox{1cm}{$\boldsymbol{v}$}]
	\item [\parbox{1cm}{$\boldsymbol{w}_{t}^{2}$}]
	\item [\parbox{1cm}{$\boldsymbol{w}_{t}$}]
	\item [\parbox{1cm}{$\boldsymbol{w}_{t+1}$}]
	\item [\parbox{1cm}{$\boldsymbol{w}$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i_{1}}(t)$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i_{1}}$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i_{2}}(t)$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i_{2}}$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i_{3}}(t)$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i_{3}}$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i} \sim U(x_{min}, x_{max})$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i}(0)$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i}(t) \in \mathcal{C}(t)$}]
	\item [\parbox{1cm}{$\boldsymbol{x}_{i}(t)$}]
	\item [\parbox{1cm}{$\boldsymbol{x}'_{i}(t)$}]
	\item [\parbox{1cm}{$\boldsymbol{y}_{i}(0)$}]
	\item [\parbox{1cm}{$\boldsymbol{y}_{i}(t)$}]
	\item [\parbox{1cm}{$\Delta w_{i}(t)$}]
	\item [\parbox{1cm}{$\Delta$}]
	\item [\parbox{1cm}{$\emptyset$}]
	\item [\parbox{1cm}{$\epsilon_{T_{p}}$}]
	\item [\parbox{1cm}{$\epsilon_{T}$}]
	\item [\parbox{1cm}{$\epsilon$}]
	\item [\parbox{1cm}{$\eta$}]
	\item [\parbox{1cm}{$\hat{\boldsymbol{m}}_{t}$}]
	\item [\parbox{1cm}{$\hat{\boldsymbol{v}}_{t}$}]
	\item [\parbox{1cm}{$\hat{y}_{ij}(t)$}]
	\item [\parbox{1cm}{$\hat{y}_{ij}$}]
	\item [\parbox{1cm}{$\hat{y}$}]
	\item [\parbox{1cm}{$\mathcal{C}(0)$}]
	\item [\parbox{1cm}{$\mathcal{C}(t+1)$}]
	\item [\parbox{1cm}{$\mathcal{E}$}]
	\item [\parbox{1cm}{$\mathcal{J}$}]
	\item [\parbox{1cm}{$\odot$}]
	\item [\parbox{1cm}{$c_{1}$}]
	\item [\parbox{1cm}{$c_{2}$}]
	\item [\parbox{1cm}{$E[\boldsymbol{g}^{2}]_{t - 1}$}]
	\item [\parbox{1cm}{$E[\boldsymbol{g}^{2}]_{t}$}]
	\item [\parbox{1cm}{$E[\Delta \boldsymbol{w}^{2}]_{t - 1}$}]
	\item [\parbox{1cm}{$E[\Delta \boldsymbol{w}^{2}]_{t}$}]
	\item [\parbox{1cm}{$f(\boldsymbol{x}_{i}(t))$}]
	\item [\parbox{1cm}{$f(\boldsymbol{x}'_{i}(t))$}]
	\item [\parbox{1cm}{$f(\hat{y})$}]
	\item [\parbox{1cm}{$f(x_{i})$}]
	\item [\parbox{1cm}{$f(x)$}]
	\item [\parbox{1cm}{$f(y_{i})$}]
	\item [\parbox{1cm}{$f$}]
	\item [\parbox{1cm}{$g_{t,i}$}]
	\item [\parbox{1cm}{$G_{t,i}$}]
	\item [\parbox{1cm}{$G_{t,ii}$}]
	\item [\parbox{1cm}{$g$}]
	\item [\parbox{1cm}{$i_{1}$}]
	\item [\parbox{1cm}{$i_{2}$}]
	\item [\parbox{1cm}{$i_{3}$}]
	\item [\parbox{1cm}{$i$}]
	\item [\parbox{1cm}{$I$}]
	\item [\parbox{1cm}{$j^{*}$}]
	\item [\parbox{1cm}{$j$}]
	\item [\parbox{1cm}{$J$}]
	\item [\parbox{1cm}{$k$}]
	\item [\parbox{1cm}{$m_{j}(t)$}]
	\item [\parbox{1cm}{$mathcal{J}$}]
	\item [\parbox{1cm}{$N_{t}$}]
	\item [\parbox{1cm}{$N$}]
	\item [\parbox{1cm}{$net_{p}$}]
	\item [\parbox{1cm}{$o_{k,p}^{2}$}]
	\item [\parbox{1cm}{$o_{p}$}]
	\item [\parbox{1cm}{$o_p$}]
	\item [\parbox{1cm}{$p_{c}$}]
	\item [\parbox{1cm}{$p_{m}$}]
	\item [\parbox{1cm}{$p_{r}$}]
	\item [\parbox{1cm}{$p_{x}$}]
	\item [\parbox{1cm}{$p$}]
	\item [\parbox{1cm}{$r_{1_{j}}$}]
	\item [\parbox{1cm}{$r_{2_{j}}$}]
	\item [\parbox{1cm}{$RMS$}]
	\item [\parbox{1cm}{$t_{k,p}$}]
	\item [\parbox{1cm}{$t_{p}$}]
	\item [\parbox{1cm}{$t$}]
	\item [\parbox{1cm}{$T$}]
	\item [\parbox{1cm}{$u_{ij}(t)$}]
	\item [\parbox{1cm}{$v_{ij}(t)$}]
	\item [\parbox{1cm}{$v_{ij}(t+1)$}]
	\item [\parbox{1cm}{$v_{ij}$}]
	\item [\parbox{1cm}{$V_{max,j}$}]
	\item [\parbox{1cm}{$V_{max}$}]
	\item [\parbox{1cm}{$v'_{ij}(t+1)$}]
	\item [\parbox{1cm}{$w_{i}(t-1)$}]
	\item [\parbox{1cm}{$w_{i}(t)$}]
	\item [\parbox{1cm}{$w_{i}$}]
	\item [\parbox{1cm}{$w_{i}$}]
	\item [\parbox{1cm}{$w_{t,i}$}]
	\item [\parbox{1cm}{$w_{t}$}]
	\item [\parbox{1cm}{$w_{t+1,i}$}]
	\item [\parbox{1cm}{$w_{t+1}$}]
	\item [\parbox{1cm}{$w$}]
	\item [\parbox{1cm}{$W$}]
	\item [\parbox{1cm}{$wv_{ij}(t)$}]
	\item [\parbox{1cm}{$wv_{ij}$}]
	\item [\parbox{1cm}{$x_{i}$}]
	\item [\parbox{1cm}{$x_{ij}(t)$}]
	\item [\parbox{1cm}{$x_{ij}(t+1)$}]
	\item [\parbox{1cm}{$x_{ij}$}]
	\item [\parbox{1cm}{$x_{max,j}$}]
	\item [\parbox{1cm}{$x_{min,j}$}]
	\item [\parbox{1cm}{$x'_{ij}(t)$}]
	\item [\parbox{1cm}{$x$}]
	\item [\parbox{1cm}{$y_{i}$}]
	\item [\parbox{1cm}{$y_{ij}(t)$}]
	\item [\parbox{1cm}{$y_{ij}$}]
	\item [\parbox{1cm}{$y$}]
	\item [\parbox{1cm}{$z_{i,p}$}]
	\item [\parbox{1cm}{$z$}]
\end{description}

\section{Chapter~\ref{chap:hhs}: Hyper-Heuristics}
\label{sec:symbols:hhs}

\begin{description}
	\item [\parbox{1cm}{$K$}]
\end{description}

\section{Chapter~\ref{chap:probability}: Probability}
\label{sec:symbols:probability}

\begin{description}
	\item [\parbox{1cm}{$\alpha_{0}$}]
	\item [\parbox{1cm}{$\alpha_{1}$}]
	\item [\parbox{1cm}{$\alpha_{k}$}]
	\item [\parbox{1cm}{$\alpha_{K}$}]
	\item [\parbox{1cm}{$\alpha$}]
	\item [\parbox{1cm}{$\bar{A}$}]
	\item [\parbox{1cm}{$\beta'$ }]
	\item [\parbox{1cm}{$\beta$}]
	\item [\parbox{1cm}{$Bin$}]
	\item [\parbox{1cm}{$\boldsymbol{\alpha'}$}]
	\item [\parbox{1cm}{$\boldsymbol{\alpha}$}]
	\item [\parbox{1cm}{$\boldsymbol{\mathcal{D}}$}]
	\item [\parbox{1cm}{$\boldsymbol{\theta}$}]
	\item [\parbox{1cm}{$\boldsymbol{N}$}]
	\item [\parbox{1cm}{$\boldsymbol{x_{1}}$}]
	\item [\parbox{1cm}{$\boldsymbol{x_{i}}$}]
	\item [\parbox{1cm}{$\boldsymbol{x_{I}}$}]
	\item [\parbox{1cm}{$\boldsymbol{x}$}]
	\item [\parbox{1cm}{$\boldsymbol{X}$}]
	\item [\parbox{1cm}{$\boldsymbol{Y}$}]
	\item [\parbox{1cm}{$\emptyset$}]
	\item [\parbox{1cm}{$\Gamma'(n)$}]
	\item [\parbox{1cm}{$\Gamma(n)$}]
	\item [\parbox{1cm}{$\Gamma$}]
	\item [\parbox{1cm}{$\mathbb{R}$}]
	\item [\parbox{1cm}{$\mathbbm{1}$}]
	\item [\parbox{1cm}{$\mathcal{A}$}]
	\item [\parbox{1cm}{$\mathcal{L}$}]
	\item [\parbox{1cm}{$\overset{\text{iid}}{\sim}$}]
	\item [\parbox{1cm}{$\psi$}]
	\item [\parbox{1cm}{$\theta_{1}$}]
	\item [\parbox{1cm}{$\theta_{k}$}]
	\item [\parbox{1cm}{$\theta_{K}$}]
	\item [\parbox{1cm}{$\theta$}]
	\item [\parbox{1cm}{$A$}]
	\item [\parbox{1cm}{$B_{1}$}]
	\item [\parbox{1cm}{$B_{2}$}]
	\item [\parbox{1cm}{$B_{i}$}]
	\item [\parbox{1cm}{$B_{j}$}]
	\item [\parbox{1cm}{$B_{K}$}]
	\item [\parbox{1cm}{$B(\boldsymbol{\alpha})$}]
	\item [\parbox{1cm}{$B$}]
	\item [\parbox{1cm}{$Ber(\theta)$}]
	\item [\parbox{1cm}{$Beta$}]
	\item [\parbox{1cm}{$Bin$}]
	\item [\parbox{1cm}{$c$}]
	\item [\parbox{1cm}{$Cat$}]
	\item [\parbox{1cm}{$d$}]
	\item [\parbox{1cm}{$Dir$}]
	\item [\parbox{1cm}{$dn$}]
	\item [\parbox{1cm}{$E[{x_{k}}]$}]
	\item [\parbox{1cm}{$E[\ln(x_{k})]$}]
	\item [\parbox{1cm}{$E[\ln(x)]$}]
	\item [\parbox{1cm}{$E[x_{k}]$}]
	\item [\parbox{1cm}{$E[x]$}]
	\item [\parbox{1cm}{$f_{Ber}$}]
	\item [\parbox{1cm}{$f_{Beta}$}]
	\item [\parbox{1cm}{$f_{Bin}$}]
	\item [\parbox{1cm}{$f_{Cat}$}]
	\item [\parbox{1cm}{$f_{Dir}$}]
	\item [\parbox{1cm}{$f_{Mul}$}]
	\item [\parbox{1cm}{$f$}]
	\item [\parbox{1cm}{$F$}]
	\item [\parbox{1cm}{$g^{*}$}]
	\item [\parbox{1cm}{$g$}]
	\item [\parbox{1cm}{$i$}]
	\item [\parbox{1cm}{$I$}]
	\item [\parbox{1cm}{$j$}]
	\item [\parbox{1cm}{$K$}]
	\item [\parbox{1cm}{$k$}]
	\item [\parbox{1cm}{$K$}]
	\item [\parbox{1cm}{$M[x_{k}]$}]
	\item [\parbox{1cm}{$M[x]$}]
	\item [\parbox{1cm}{$m$}]
	\item [\parbox{1cm}{$M$}]
	\item [\parbox{1cm}{$Mul$}]
	\item [\parbox{1cm}{$N_{0}$}]
	\item [\parbox{1cm}{$N_{1}$}]
	\item [\parbox{1cm}{$N_{k}$}]
	\item [\parbox{1cm}{$N_{K}$}]
	\item [\parbox{1cm}{$n$}]
	\item [\parbox{1cm}{$N$}]
	\item [\parbox{1cm}{$p$}]
	\item [\parbox{1cm}{$S$}]
	\item [\parbox{1cm}{$v$}]
	\item [\parbox{1cm}{$x_{1}$}]
	\item [\parbox{1cm}{$x_{i,k}$}]
	\item [\parbox{1cm}{$x_{i}$}]
	\item [\parbox{1cm}{$x_{I}$}]
	\item [\parbox{1cm}{$x_{j}$}]
	\item [\parbox{1cm}{$x_{k}$}]
	\item [\parbox{1cm}{$x_{K}$}]
	\item [\parbox{1cm}{$x$}]
	\item [\parbox{1cm}{$y_{1}$}]
	\item [\parbox{1cm}{$Y_{1}$}]
	\item [\parbox{1cm}{$y_{2}$}]
	\item [\parbox{1cm}{$Y_{2}$}]
	\item [\parbox{1cm}{$y_{i}$}]
	\item [\parbox{1cm}{$Y_{i}$}]
	\item [\parbox{1cm}{$y_{n}$}]
	\item [\parbox{1cm}{$Y_{N}$}]
\end{description}

\section{Chapter~\ref{chap:bhh}: Bayesian Hyper-Heuristic}
\label{sec:symbols:bhh}

\begin{description}
	\item [\parbox{1cm}{$\alpha_{1}$}]
	\item [\parbox{1cm}{$\alpha_{k}$}]
	\item [\parbox{1cm}{$\alpha_{K}$}]
	\item [\parbox{1cm}{$\alpha$}]
	\item [\parbox{1cm}{$\beta_{1}$}]
	\item [\parbox{1cm}{$\beta_{2}$}]
	\item [\parbox{1cm}{$\beta_{j,k}$}]
	\item [\parbox{1cm}{$\beta_{K}^{J}$}]
	\item [\parbox{1cm}{$\beta$}]
	\item [\parbox{1cm}{$\boldsymbol{\alpha}$}]
	\item [\parbox{1cm}{$\boldsymbol{\beta}$}]
	\item [\parbox{1cm}{$\boldsymbol{\gamma}$}]
	\item [\parbox{1cm}{$\boldsymbol{\hat{\phi}}$}]
	\item [\parbox{1cm}{$\boldsymbol{\hat{\psi}}$}]
	\item [\parbox{1cm}{$\boldsymbol{\hat{\theta}}$}]
	\item [\parbox{1cm}{$\boldsymbol{\lambda}$}]
	\item [\parbox{1cm}{$\boldsymbol{\phi}$}]
	\item [\parbox{1cm}{$\boldsymbol{\psi}$}]
	\item [\parbox{1cm}{$\boldsymbol{\theta}$}]
	\item [\parbox{1cm}{$\boldsymbol{C}$}]
	\item [\parbox{1cm}{$\boldsymbol{E}$}]
	\item [\parbox{1cm}{$\boldsymbol{H}$}]
	\item [\parbox{1cm}{$\epsilon(-1)$}]
	\item [\parbox{1cm}{$\epsilon$}]
	\item [\parbox{1cm}{$\eta$}]
	\item [\parbox{1cm}{$\gamma_{0}$}]
	\item [\parbox{1cm}{$\gamma_{1,k}$}]
	\item [\parbox{1cm}{$\gamma_{1}$}]
	\item [\parbox{1cm}{$\gamma_{2,k}$}]
	\item [\parbox{1cm}{$\gamma_{K}$}]
	\item [\parbox{1cm}{$\hat{\phi}_{j,k}$}]
	\item [\parbox{1cm}{$\hat{\psi_{k}}$}]
	\item [\parbox{1cm}{$\hat{\psi}_{k}$}]
	\item [\parbox{1cm}{$\hat{\theta_{k}}$}]
	\item [\parbox{1cm}{$\hat{\theta}_{k}$}]
	\item [\parbox{1cm}{$\lambda_{1}$}]
	\item [\parbox{1cm}{$\lambda_{j}$}]
	\item [\parbox{1cm}{$\lambda_{J}$}]
	\item [\parbox{1cm}{$\mathbbm{1}_{0}$}]
	\item [\parbox{1cm}{$\mathbbm{1}_{1}$}]
	\item [\parbox{1cm}{$\mathbin{\%}$}]
	\item [\parbox{1cm}{$\mathcal{A}(v)$}]
	\item [\parbox{1cm}{$\mathcal{L}$}]
	\item [\parbox{1cm}{$\phi_{j,k}$}]
	\item [\parbox{1cm}{$\phi_{jk}$}]
	\item [\parbox{1cm}{$\phi$}]
	\item [\parbox{1cm}{$\psi_{k}$}]
	\item [\parbox{1cm}{$\psi_{k}N_{k}$}]
	\item [\parbox{1cm}{$\psi$}]
	\item [\parbox{1cm}{$\theta_{k}$}]
	\item [\parbox{1cm}{$\theta$}]
	\item [\parbox{1cm}{$A$}]
	\item [\parbox{1cm}{$B$}]
	\item [\parbox{1cm}{$Beta$}]
	\item [\parbox{1cm}{$Bin$}]
	\item [\parbox{1cm}{$c_{1}$}]
	\item [\parbox{1cm}{$c_{i,k}$}]
	\item [\parbox{1cm}{$c_{i}$}]
	\item [\parbox{1cm}{$Dir$}]
	\item [\parbox{1cm}{$e_{i,j,k}$}]
	\item [\parbox{1cm}{$e_{i}$}]
	\item [\parbox{1cm}{$e_{j}$}]
	\item [\parbox{1cm}{$E[\phi_{j,k}]$}]
	\item [\parbox{1cm}{$E[\psi_{k}]$}]
	\item [\parbox{1cm}{$E[\theta_{k}]$}]
	\item [\parbox{1cm}{$h_{i,k}$}]
	\item [\parbox{1cm}{$h_{i}$}]
	\item [\parbox{1cm}{$h_{k}$}]
	\item [\parbox{1cm}{$i$}]
	\item [\parbox{1cm}{$I$}]
	\item [\parbox{1cm}{$j$}]
	\item [\parbox{1cm}{$J$}]
	\item [\parbox{1cm}{$K-1$}]
	\item [\parbox{1cm}{$k$}]
	\item [\parbox{1cm}{$K$}]
	\item [\parbox{1cm}{$L$}]
	\item [\parbox{1cm}{$LSE$}]
	\item [\parbox{1cm}{$Mult$}]
	\item [\parbox{1cm}{$N_{0,k}$}]
	\item [\parbox{1cm}{$N_{1,k}$}]
	\item [\parbox{1cm}{$N_{j,k}$}]
	\item [\parbox{1cm}{$N_{j}$}]
	\item [\parbox{1cm}{$N_{k}$}]
	\item [\parbox{1cm}{$N$}]
	\item [\parbox{1cm}{$S$}]
	\item [\parbox{1cm}{$t$}]
\end{description}

\section{Chapter~\ref{chap:methodology}: Methodology}
\label{sec:symbols:methodology}

\begin{description}
	\item [\parbox{1cm}{$\alpha$}]
	\item [\parbox{1cm}{$\eta$}]
	\item [\parbox{1cm}{$\mathbb{R}$}]
	\item [\parbox{1cm}{$v_{ij}$}]
	\item [\parbox{1cm}{$x_{ij}$}]

\end{description}

\section{Chapter~\ref{chap:results}: Results}
\label{sec:symbols:results}

\begin{description}
	\item [\parbox{1cm}{$\alpha_{0}$}]
	\item [\parbox{1cm}{$\alpha_{6}$}]
	\item [\parbox{1cm}{$\alpha_{7}$}]
	\item [\parbox{1cm}{$\alpha_{8}$}]
	\item [\parbox{1cm}{$\alpha_{i} = 50$}]
	\item [\parbox{1cm}{$\boldsymbol{{H}}$}]
	\item [\parbox{1cm}{$\boldsymbol{\alpha}$}]
	\item [\parbox{1cm}{$\boldsymbol{\phi}$}]
	\item [\parbox{1cm}{$\boldsymbol{\psi}$}]
	\item [\parbox{1cm}{$\boldsymbol{\theta}$}]
	\item [\parbox{1cm}{$\boldsymbol{C}$}]
	\item [\parbox{1cm}{$\boldsymbol{E}$}]
	\item [\parbox{1cm}{$\boldsymbol{H}$}]
	\item [\parbox{1cm}{$\pm$}]
	\item [\parbox{1cm}{$\theta_{0}$}]
	\item [\parbox{1cm}{$\theta_{6}$}]
	\item [\parbox{1cm}{$\theta_{7}$}]
	\item [\parbox{1cm}{$\theta_{8}$}]
	\item [\parbox{1cm}{$c_{1}$}]
	\item [\parbox{1cm}{$Cat$}]
	\item [\parbox{1cm}{$Dir$}]
	\item [\parbox{1cm}{$e_{0}$}]
	\item [\parbox{1cm}{$h_{0}$}]
	\item [\parbox{1cm}{$h_{6}$}]
	\item [\parbox{1cm}{$h_{7}$}]
	\item [\parbox{1cm}{$h_{8}$}]
	\item [\parbox{1cm}{$i$}]
	\item [\parbox{1cm}{$K$}]
\end{description}