### 2022-12-09 dissertation/appendix/symbols.tex
✨ (symbols): Chapters 3,4,7 done

\chapter{Symbols}
\label{app:symbols}

This appendix summarises the important symbols used throughout the dissertation. For each symbol, a short description is provided as well as the reference to the equation that makes use it. Symbols are divided according to the chapter in which they were introduced and are ordered alphabetically.

\section{Chapter~\ref{chap:anns}: Artificial Neural Networks}
\label{sec:symbols:anns}

\begin{description}
	% \item
	%       [\parbox{2cm}{$EXAMPLE$}] THIS IS AN EXAMPLE DESCRIPTION. \lbrack Eq.~\eqref{eq:an_function_mapping} pg.~\pageref{eq:an_function_mapping} \rbrack

	\item [\parbox{2cm}{$I$}] The input pattern's dimension size.
	\item [\parbox{2cm}{$T$}] The target pattern's dimension size.
	\item [\parbox{2cm}{$\mathbb{R}$}] The real-number space.
	\item [\parbox{2cm}{$\mathbb{R}^{I}$}] The real-number space in I dimensions.
	\item [\parbox{2cm}{$\mathbb{R}^{T}$}] The real-number space in T dimensions.
	\item [\parbox{2cm}{$f_{AN}$}] The mapping function realised by the \acs{AN}.
	\item [\parbox{2cm}{$\boldsymbol{x}$}] The input pattern/vector.
	\item [\parbox{2cm}{$x_{i}$}] The $i$-th dimension of the input pattern/vector.
	\item [\parbox{2cm}{$\theta$}] The bias unit.
	\item [\parbox{2cm}{$\boldsymbol{v}$}] The weight vector.
	\item [\parbox{2cm}{$v_{i}$}] The $i$-th dimension of the weight vector.
	\item [\parbox{2cm}{$net$}] The net input signal to the \acs{AN}.
	\item [\parbox{2cm}{$f(net)$}] The activation function over net input signal.
	\item [\parbox{2cm}{$\boldsymbol{y}$}] The output vector.
	\item [\parbox{2cm}{$\boldsymbol{y}_{p}$}] The output vector for pattern $p$.
	\item [\parbox{2cm}{$k$}] Index for the output vector.
	\item [\parbox{2cm}{$x_{i,p}^{'}$}] The normalised form of the input at dimension $i$, pattern $p$.
	\item [\parbox{2cm}{$x_{i,p}$}] The original input at dimension $i$, pattern $p$.
	\item [\parbox{2cm}{$x_{i_{min}}$}] The input's minimum value at dimension $i$.
	\item [\parbox{2cm}{$x_{i_{max}}$}] The input's maximum value at dimension $i$.
	\item [\parbox{2cm}{$p$}] Index for the input data patterns.
	\item [\parbox{2cm}{$P$}] Total number of input data patterns.
	\item [\parbox{2cm}{$\boldsymbol{x}_p$}] The $p$-th input vector/pattern.
	\item [\parbox{2cm}{$\mu_i$}] The mean of the input at dimension $i$.
	\item [\parbox{2cm}{$\sigma^2_i$}] The unit variance of the input at dimension $i$.
	\item [\parbox{2cm}{$i$}] Index for the input vector.
	\item [\parbox{2cm}{$\omega$}] The general weight vector used for normalisation.
	\item [\parbox{2cm}{$\omega_{i}$}] The general weight vector used for normalisation at dimension $i$.
	\item [\parbox{2cm}{$\omega_{min}$}] The lower bound of the uniform distribution used for weights initialisation.
	\item [\parbox{2cm}{$\omega_{max}$}] The upper bound of the uniform distribution used for weights initialisation.
	\item [\parbox{2cm}{$fanin$}] The number of input neurons to the weight vector.
	\item [\parbox{2cm}{$fanout$}] The number of output neurons from the weight vector.
	\item [\parbox{2cm}{$\sigma$}] The standard deviation of the truncated normal distribution used for weight initialisation.
	\item [\parbox{2cm}{$K$}] Number of output units.
	\item [\parbox{2cm}{$x_{i+1}$}] The $i+1$-th dimension of the input pattern/vector.
	\item [\parbox{2cm}{$v_{i+1}$}] The $i+1$-th dimension of the weight vector.
	\item [\parbox{2cm}{$net^{'}$}] The augmented net input signal that includes the bias term.
	\item [\parbox{2cm}{$\tau$}] Phase shift threshold parameter.
	\item [\parbox{2cm}{$f(x)$}] The function used to describe the \acs{LReLU} activation function.
	\item [\parbox{2cm}{$f$}] A shorted form of the activation function.
	\item [\parbox{2cm}{$x^{+}$}] The positive part of the input parameter's domain.
	\item [\parbox{2cm}{$\alpha$}] Scaling parameter used by the \acs{LReLU} activation function.
	\item [\parbox{2cm}{$\lambda$}] Scaling/control parameter used by the sigmoid and tanh activation functions.
	\item [\parbox{2cm}{$x_{min}$}] The input's minimum value used by the min-max scaler.
	\item [\parbox{2cm}{$x_{max}$}] The input's maximum value used by the min-max scaler.
	\item [\parbox{2cm}{$\mathbb{R}^{K}$}] The real-number space in $K$ dimensions.
	\item [\parbox{2cm}{$y_{k}$}] The $k$-th dimension of the output vector.
	\item [\parbox{2cm}{$y^{'}_k$}] The softmax value of the output at dimension $k$.
	\item [\parbox{2cm}{$\boldsymbol{y^{'}}$}] The softmax form of the output vector.
	\item [\parbox{2cm}{$h_{j}$}] The $j$-th dimension of the hidden layer.
	\item [\parbox{2cm}{$j$}] Index used for the hidden layer.
	\item [\parbox{2cm}{$v_{i,j}$}] The weight associated with input node $x_{i}$ and the hidden node $h_{j}$.
	\item [\parbox{2cm}{$w_{j,k}$}] The weight associated with hidden node $h_{j}$ and the output node $y_{k}$.
	\item [\parbox{2cm}{$net_{i,h}$}] The net input signal at the hidden layer.
	\item [\parbox{2cm}{$net_{h,y}$}] The net input signal at the output layer.
	\item [\parbox{2cm}{$\epsilon$}] The cost/loss produced by the loss function.
	\item [\parbox{2cm}{$y_{k,p}$}] The output produced by the \acs{ANN} at the $k$-th dimension for the $p$-th pattern.
	\item [\parbox{2cm}{$\hat{y}_{k,p}$}] The target output for the \acs{ANN} at the $k$-th dimension for the $p$-th pattern.
	\item [\parbox{2cm}{$\beta$}] Mini-batch size.
	\item [\parbox{2cm}{$\mathbbm{1}$}] The indicator function.
	\item [\parbox{2cm}{$c$}] Index used for $c$-th class.
	\item [\parbox{2cm}{$C$}] Total number of classes.

\end{description}

\section{Chapter~\ref{chap:heuristics}: Heuristics}
\label{sec:symbols:heuristics}

\begin{description}
	\item [\parbox{2cm}{$f$}] A shortened form of the objective function being optimised.
	\item [\parbox{2cm}{$x$}] The independent variables to the objective function. Also used in \acs{DE} notation to denote the selection mechanism for the trial vector.
	\item [\parbox{2cm}{$f(x)$}] The objective function over $x$.
	\item [\parbox{2cm}{$\epsilon$}] The error between the output of the \acs{FFNN} and the target pattern, produced by the loss function. Also used as a small error value used by heuristics to avoid division by 0.
	\item [\parbox{2cm}{$\mathcal{E}$}] The error function.
	\item [\parbox{2cm}{$w_{i}(t)$}] The $i$-th dimension of the weight vector at time step $t$.
	\item [\parbox{2cm}{$w_{i}(t-1)$}] The $i$-th dimension of the weight vector at time step $t-1$.
	\item [\parbox{2cm}{$\Delta w_{i}(t)$}] The change in the $i$-th dimension of the weight vector at time step $t$.
	\item [\parbox{2cm}{$w_{i}$}] The $i$-th dimension of the weight vector.
	\item [\parbox{2cm}{$i$}] Index used for the weight vector. Also used as the index for the $i$-th particle in the population, implemented by \acs{PSO}.
	\item [\parbox{2cm}{$\boldsymbol{w}$}] The weight vector.
	\item [\parbox{2cm}{$p$}] Index used for input data patterns.
	\item [\parbox{2cm}{$o_{i,p}$}] The $i$-th dimension of the output for pattern $p$.
	\item [\parbox{2cm}{$t_{i,p}$}] The $i$-th dimension of the target for pattern $p$.
	      item [\parbox{2cm}{$o_{k,p}$}] The $k$-th dimension of the output for pattern $p$.
	\item [\parbox{2cm}{$t_{k,p}$}] The $k$-th dimension of the target for pattern $p$.
	\item [\parbox{2cm}{$net_{p}$}] The net input signal for pattern $p$.
	\item [\parbox{2cm}{$z_{i,p}$}] The input value at index $i$ for pattern $p$.
	\item [\parbox{2cm}{$\boldsymbol{g}$}] General symbol used to represent the gradient vector, retrieved from the error function relative to the weight vector.
	\item [\parbox{2cm}{$\boldsymbol{v}$}] The velocity vector resulting from \acs{Momentum}.
	\item [\parbox{2cm}{$\boldsymbol{w}$}] General symbol used to represent the weight vector.
	\item [\parbox{2cm}{$\epsilon_{T}$}] The error at time step $T$.
	\item [\parbox{2cm}{$\epsilon_{T_{p}}$}] The error produced for pattern $p$ at time step $T$.
	\item [\parbox{2cm}{$\eta$}] Learning rate.
	\item [\parbox{2cm}{$k$}] Index used for the output layer.
	\item [\parbox{2cm}{$K$}] Total number of output units.
	\item [\parbox{2cm}{$t$}] Time step.
	\item [\parbox{2cm}{$T$}] Time step.
	\item [\parbox{2cm}{$\alpha$}] The momentum hyper-parameter.
	\item [\parbox{2cm}{$w_{t,i}$}] The $i$-th dimension of the weight vector at time step $t$.
	\item [\parbox{2cm}{$w_{t+1,i}$}] The $i$-th dimension of the weight vector at time step $t+1$.
	\item [\parbox{2cm}{$G_{t,ii}$}] Diagonal matrix containing the sum of the squared gradients with regards to the $i$-th dimension of the weight vector.
	\item [\parbox{2cm}{$g_{t,i}$}] The $i$-th dimension of the gradient vector at time step $t$.
	\item [\parbox{2cm}{$\boldsymbol{G_{t}}$}] The sum of the squares of the gradients.
	\item [\parbox{2cm}{$G_{t,i}$}] The sum of the squared gradients with regards to the $i$-th dimension of the weight vector.
	\item [\parbox{2cm}{$\odot$}] The matrix-vector product.
	\item [\parbox{2cm}{$W$}] Window size of previous squared gradients.
	\item [\parbox{2cm}{$E[\boldsymbol{g}^{2}]_{t}$}] The expected value of the squared gradients at time step $t$.
	\item [\parbox{2cm}{$E[\boldsymbol{g}^{2}]_{t - 1}$}] The expected value of the squared gradients at time step $t-1$.
	\item [\parbox{2cm}{$\boldsymbol{g}^{2}_{t}$}] The squared gradient vector at time step $t$.
	\item [\parbox{2cm}{$\boldsymbol{w}_{t+1}$}] The weight vector at time step $t+1$.
	\item [\parbox{2cm}{$\boldsymbol{w}_{t}^{2}$}] The squared weight vector at time step $t$.
	\item [\parbox{2cm}{$\boldsymbol{g}_{t}$}] The gradient vector at time step $t$.
	\item [\parbox{2cm}{$RMS$}] The root mean squared error criterion.
	\item [\parbox{2cm}{$E[\Delta \boldsymbol{w}^{2}]_{t}$}] The expected value of the delta of the squared weight vector at time step $t$.
	\item [\parbox{2cm}{$E[\Delta \boldsymbol{w}^{2}]_{t - 1}$}] The expected value of the delta of the squared weight vector at time step $t-1$.
	\item [\parbox{2cm}{$\Delta \boldsymbol{w}^{2}_{t}$}] The delta of the squared weight vector at time step $t$.
	\item [\parbox{2cm}{$RMS[\Delta \boldsymbol{w}]_{t}$}] The root mean squared delta of the weight vector at time step $t$.
	\item [\parbox{2cm}{$RMS[\Delta \boldsymbol{w}]_{t-1}$}] The root mean squared delta of the weight vector at time step $t-1$.
	\item [\parbox{2cm}{$RMS[\boldsymbol{g}]_{t}$}] The root mean squared gradient vector at time step $t$.
	\item [\parbox{2cm}{$\boldsymbol{g}_{t}$}] The gradient vector at time step $t$.
	\item [\parbox{2cm}{$\boldsymbol{m}_{t}$}] The estimate of the first moment at time step $t$.
	\item [\parbox{2cm}{$\boldsymbol{m}_{t+1}$}] The estimate of the first moment at time step $t+1$.
	\item [\parbox{2cm}{$\boldsymbol{v}_{t}$}] The estimate of the second moment at time step $t$.
	\item [\parbox{2cm}{$\boldsymbol{v}_{t+1}$}] The estimate of the second moment at time step $t+1$.
	\item [\parbox{2cm}{$\beta_{1}$}] First decay rate parameter for \acs{Adam}.
	\item [\parbox{2cm}{$\beta_{2}$}] Second decay rate parameter for \acs{Adam}.
	\item [\parbox{2cm}{$\hat{\boldsymbol{m}}_{t}$}] The bias-corrected first moment at time step $t$.
	\item [\parbox{2cm}{$\hat{\boldsymbol{v}}_{t}$}] The bias-corrected second moment at time step $t$.
	\item [\parbox{2cm}{$\beta^{t}_{1}$}] First decay rate parameter for \acs{Adam} at time step $t$.
	\item [\parbox{2cm}{$\beta^{t}_{2}$}] Second decay rate parameter for \acs{Adam} at time step $t$.
	\item [\parbox{2cm}{$\boldsymbol{x}_{i}$}] The $i$-th particle's position vector as implemented by \acs{PSO}. Also represents the $i$-th parent's position vector as implemented by \acs{DE}.
	\item [\parbox{2cm}{$\boldsymbol{x}_{i}(t)$}] The $i$-th particle's position vector at time step $t$ as implemented by \acs{PSO}. Also represents the $i$-th parent's position vector at time step $t$ as implemented by \acs{DE}.
	\item [\parbox{2cm}{$\boldsymbol{v}_{i}(t)$}] The $i$-th particle's velocity vector at time step $t$ as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$f(x_{i})$}] The evaluation of the $i$-th particle's solution with respect to the objective function.
	\item [\parbox{2cm}{$\boldsymbol{y}_{i}(t)$}] The $i$-th particle's personal best position vector at time step $t$ as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$\boldsymbol{x}_{i}(0)$}] The $i$-th particle's position vector at time step $0$ as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$\boldsymbol{y}_{i}(0)$}] The $i$-th particle's personal best position vector at time step $0$ as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$\boldsymbol{\hat{y}}_{i}(t)$}] The $i$-th particle's global best position vector as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$\boldsymbol{x}_{i}$}] The $i$-th particle's position vector as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$\boldsymbol{x}_{min}$}] The lower bound constraint vector of the position vector's values.
	\item [\parbox{2cm}{$\boldsymbol{x}_{max}$}] The upper bound constraint vector of the position vector's values.
	\item [\parbox{2cm}{$v_{ij}(t+1)$}] The $i$-th particle's $j$-th dimension of the velocity vector at time step $t+1$.
	\item [\parbox{2cm}{$v_{ij}(t)$}] The $i$-th particle's $j$-th dimension of the velocity vector at time step $t$. Also used in the mutation operator for \acp{GA} to represent some sampled mutation update value.
	\item [\parbox{2cm}{$w$}] The inertia weight hyper-parameter as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$c_{1}$}] The cognitive control parameter as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$c_{2}$}] The social control parameter as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$r_{1_{j}}(t)$}] The $j$-th dimension of the first stochastic scaling parameter at time step $t$ as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$r_{2_{j}}(t)$}] The $j$-th dimension of the second stochastic scaling parameter at time step $t$ as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$x_{ij}$}] The $i$-th particle's/entity's $j$-th dimension of the position vector.
	\item [\parbox{2cm}{$x_{ij}(t)$}] The $i$-th particle's/entity's $j$-th dimension of the position vector at time step $t$.
	\item [\parbox{2cm}{$y_{ij}$}] The $i$-th particle's $j$-th dimension of the personal best position vector.
	\item [\parbox{2cm}{$y_{ij}(t)$}] The $i$-th particle's $j$-th dimension of the personal best position vector at time step $t$.
	\item [\parbox{2cm}{$\boldsymbol{\hat{y}}$}] Short form of the global best position vector.
	\item [\parbox{2cm}{$\hat{y}_{ij}$}] The $i$-th particle's $j$-th dimension of the global best position vector.
	\item [\parbox{2cm}{$\hat{y}_{ij}(t)$}] The $i$-th particle's $j$-th dimension of the global best position vector at time step $t$.
	\item [\parbox{2cm}{$f(\hat{y})$}] The evaluation of the global best position.
	\item [\parbox{2cm}{$x_{ij}(t+1)$}] The $i$-th particle's $j$-th dimension of the position vector at time step $t+1$.
	\item [\parbox{2cm}{$\boldsymbol{r}_{1}$}] The vector for the first stochastic scaling parameter as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$\boldsymbol{r}_{2}$}] The vector for the second stochastic scaling parameter as implemented by \acs{PSO}.
	\item [\parbox{2cm}{$I$}] The total number of particles in the swarm (particle swarm size).
	\item [\parbox{2cm}{$N$}] An alternative to the particle swarm size.
	\item [\parbox{2cm}{$j$}] Index used for the dimensions of candidate solutions.
	\item [\parbox{2cm}{$J$}] The total number of dimensions in the position vector for particles in \acs{PSO}.
	\item [\parbox{2cm}{$U$}] The uniform distribution.
	\item [\parbox{2cm}{$v'_{ij}(t+1)$}] The $i$-th particle's $j$-th dimension of the clamped velocity vector at time step $t+1$.
	\item [\parbox{2cm}{$\boldsymbol{V}_{max}$}] The upper bound of the velocity clamping vector.
	\item [\parbox{2cm}{$V_{max,j}$}] The $j$-th dimension of the upper bound of the velocity clamping vector.
	\item [\parbox{2cm}{$\boldsymbol{u}_{i}(t)$}] The $i$-th entity's trial vector at time step $t$ implemented by \acs{DE}.
	\item [\parbox{2cm}{$\boldsymbol{x}_{i_{1}}(t)$}] The $i$-th entity's target vector at time step $t$ implemented by \acs{DE}.
	\item [\parbox{2cm}{$\boldsymbol{x}_{i_{2}}(t)$}] The $i$-th entity's first selected individual's position vector at time step $t$ implemented by \acs{DE}.
	\item [\parbox{2cm}{$\boldsymbol{x}_{i_{3}}(t)$}] The $i$-th entity's second selected individual's position vector at time step $t$ implemented by \acs{DE}.
	\item [\parbox{2cm}{$i_{1}$}] Index for the target vector in \acs{DE}.
	\item [\parbox{2cm}{$i_{2}$}] Index for the first selected individual in \acs{DE}.
	\item [\parbox{2cm}{$i_{3}$}] Index for the second selected individual in \acs{DE}.
	\item [\parbox{2cm}{$\beta$}] Scale factor as implemented by \acs{DE}.
	\item [\parbox{2cm}{$\boldsymbol{x}'_{i}(t)$}] The $i$-th offspring's position vector at time step $t$ as implemented by \acs{DE}.
	\item [\parbox{2cm}{$x'_{ij}(t)$}] The $i$-th offspring's $j$-th dimension of the position vector at time step $t$.
	\item [\parbox{2cm}{$u_{ij}(t)$}] The $j$-th dimension of the $i$-th entity's trial vector at time step $t$ implemented by \acs{DE}.
	\item [\parbox{2cm}{$\mathcal{J}$}] The set of crossover points.
	\item [\parbox{2cm}{$j^{*}$}] Randomly selected crossover point.
	\item [\parbox{2cm}{$\emptyset$}] The empty set.
	\item [\parbox{2cm}{$p_{r}$}] The crossover probability i.e. recombination probability.
	\item [\parbox{2cm}{$f(\boldsymbol{x}_{i}(t))$}] The evaluation of the $i$-th entity's candidate solution at time step $t$.
	\item [\parbox{2cm}{$f(\boldsymbol{x}'_{i}(t))$}] The evaluation of the $i$-th offspring's candidate solution at time step $t$.
	\item [\parbox{2cm}{$f(\boldsymbol{y}_{i})$}] The evaluation of the $i$-th particle's personal best position.
	\item [\parbox{2cm}{$f(\boldsymbol{\hat{y}})$}] The evaluation of the global best position.
	\item [\parbox{2cm}{$x_{min,j}$}] The $j$-th dimension of the lower bound constraint for an entity's position.
	\item [\parbox{2cm}{$x_{max,j}$}] The $j$-th dimension of the upper bound constraint for an entity's position.
	\item [\parbox{2cm}{$\mathcal{C}$}] The set of entities in the population.
	\item [\parbox{2cm}{$\mathcal{C}(0)$}] The set of entities in the population at time step $0$.
	\item [\parbox{2cm}{$\mathcal{C}(t)$}] The set of entities in the population at time step $t$.
	\item [\parbox{2cm}{$\mathcal{C}(t+1)$}] The set of entities in the population at time step $t+1$.
	\item [\parbox{2cm}{$y$}] Used in \acs{DE} notation to denote the number of difference vectors to include.
	\item [\parbox{2cm}{$z$}] Used in \acs{DE} notation to denote the type of crossover operator to use.
	\item [\parbox{2cm}{$p_{c}$}] Probability of producing offspring.
	\item [\parbox{2cm}{$p_{x}$}] Bitswapping probability.
	\item [\parbox{2cm}{$m_{j}(t)$}] The $j$-th dimension of the crossover mask, $m$, at time step $t$.
	\item [\parbox{2cm}{$p_{m}$}] The mutation probability.
	\item [\parbox{2cm}{$N_{t}$}] Tournament selection size.
\end{description}

\section{Chapter~\ref{chap:hhs}: Hyper-Heuristics}
\label{sec:symbols:hhs}

\begin{description}
	\item [\parbox{2cm}{$K$}] The number of folds to execute for $K$-fold crossover validation.
\end{description}

\section{Chapter~\ref{chap:probability}: Probability}
\label{sec:symbols:probability}

\begin{description}
	\item [\parbox{2cm}{$\alpha_{0}$}]
	\item [\parbox{2cm}{$\alpha_{1}$}]
	\item [\parbox{2cm}{$\alpha_{k}$}]
	\item [\parbox{2cm}{$\alpha_{K}$}]
	\item [\parbox{2cm}{$\alpha$}]
	\item [\parbox{2cm}{$\bar{A}$}]
	\item [\parbox{2cm}{$\beta'$ }]
	\item [\parbox{2cm}{$\beta$}]
	\item [\parbox{2cm}{$Bin$}]
	\item [\parbox{2cm}{$\boldsymbol{\alpha'}$}]
	\item [\parbox{2cm}{$\boldsymbol{\alpha}$}]
	\item [\parbox{2cm}{$\boldsymbol{\mathcal{D}}$}]
	\item [\parbox{2cm}{$\boldsymbol{\theta}$}]
	\item [\parbox{2cm}{$\boldsymbol{N}$}]
	\item [\parbox{2cm}{$\boldsymbol{x_{1}}$}]
	\item [\parbox{2cm}{$\boldsymbol{x_{i}}$}]
	\item [\parbox{2cm}{$\boldsymbol{x_{I}}$}]
	\item [\parbox{2cm}{$\boldsymbol{x}$}]
	\item [\parbox{2cm}{$\boldsymbol{X}$}]
	\item [\parbox{2cm}{$\boldsymbol{Y}$}]
	\item [\parbox{2cm}{$\emptyset$}]
	\item [\parbox{2cm}{$\Gamma'(n)$}]
	\item [\parbox{2cm}{$\Gamma(n)$}]
	\item [\parbox{2cm}{$\Gamma$}]
	\item [\parbox{2cm}{$\mathbb{R}$}]
	\item [\parbox{2cm}{$\mathbbm{1}$}]
	\item [\parbox{2cm}{$\mathcal{A}$}]
	\item [\parbox{2cm}{$\mathcal{L}$}]
	\item [\parbox{2cm}{$\overset{\text{iid}}{\sim}$}]
	\item [\parbox{2cm}{$\psi$}]
	\item [\parbox{2cm}{$\theta_{1}$}]
	\item [\parbox{2cm}{$\theta_{k}$}]
	\item [\parbox{2cm}{$\theta_{K}$}]
	\item [\parbox{2cm}{$\theta$}]
	\item [\parbox{2cm}{$A$}]
	\item [\parbox{2cm}{$B_{1}$}]
	\item [\parbox{2cm}{$B_{2}$}]
	\item [\parbox{2cm}{$B_{i}$}]
	\item [\parbox{2cm}{$B_{j}$}]
	\item [\parbox{2cm}{$B_{K}$}]
	\item [\parbox{2cm}{$B(\boldsymbol{\alpha})$}]
	\item [\parbox{2cm}{$B$}]
	\item [\parbox{2cm}{$Ber(\theta)$}]
	\item [\parbox{2cm}{$Beta$}]
	\item [\parbox{2cm}{$Bin$}]
	\item [\parbox{2cm}{$c$}]
	\item [\parbox{2cm}{$Cat$}]
	\item [\parbox{2cm}{$d$}]
	\item [\parbox{2cm}{$Dir$}]
	\item [\parbox{2cm}{$dn$}]
	\item [\parbox{2cm}{$E[{x_{k}}]$}]
	\item [\parbox{2cm}{$E[\ln(x_{k})]$}]
	\item [\parbox{2cm}{$E[\ln(x)]$}]
	\item [\parbox{2cm}{$E[x_{k}]$}]
	\item [\parbox{2cm}{$E[x]$}]
	\item [\parbox{2cm}{$f_{Ber}$}]
	\item [\parbox{2cm}{$f_{Beta}$}]
	\item [\parbox{2cm}{$f_{Bin}$}]
	\item [\parbox{2cm}{$f_{Cat}$}]
	\item [\parbox{2cm}{$f_{Dir}$}]
	\item [\parbox{2cm}{$f_{Mul}$}]
	\item [\parbox{2cm}{$f$}]
	\item [\parbox{2cm}{$F$}]
	\item [\parbox{2cm}{$g^{*}$}]
	\item [\parbox{2cm}{$g$}]
	\item [\parbox{2cm}{$i$}]
	\item [\parbox{2cm}{$I$}]
	\item [\parbox{2cm}{$j$}]
	\item [\parbox{2cm}{$K$}]
	\item [\parbox{2cm}{$k$}]
	\item [\parbox{2cm}{$K$}]
	\item [\parbox{2cm}{$M[x_{k}]$}]
	\item [\parbox{2cm}{$M[x]$}]
	\item [\parbox{2cm}{$m$}]
	\item [\parbox{2cm}{$M$}]
	\item [\parbox{2cm}{$Mul$}]
	\item [\parbox{2cm}{$N_{0}$}]
	\item [\parbox{2cm}{$N_{1}$}]
	\item [\parbox{2cm}{$N_{k}$}]
	\item [\parbox{2cm}{$N_{K}$}]
	\item [\parbox{2cm}{$n$}]
	\item [\parbox{2cm}{$N$}]
	\item [\parbox{2cm}{$p$}]
	\item [\parbox{2cm}{$S$}]
	\item [\parbox{2cm}{$v$}]
	\item [\parbox{2cm}{$x_{1}$}]
	\item [\parbox{2cm}{$x_{i,k}$}]
	\item [\parbox{2cm}{$x_{i}$}]
	\item [\parbox{2cm}{$x_{I}$}]
	\item [\parbox{2cm}{$x_{j}$}]
	\item [\parbox{2cm}{$x_{k}$}]
	\item [\parbox{2cm}{$x_{K}$}]
	\item [\parbox{2cm}{$x$}]
	\item [\parbox{2cm}{$y_{1}$}]
	\item [\parbox{2cm}{$Y_{1}$}]
	\item [\parbox{2cm}{$y_{2}$}]
	\item [\parbox{2cm}{$Y_{2}$}]
	\item [\parbox{2cm}{$y_{i}$}]
	\item [\parbox{2cm}{$Y_{i}$}]
	\item [\parbox{2cm}{$y_{n}$}]
	\item [\parbox{2cm}{$Y_{N}$}]
\end{description}

\section{Chapter~\ref{chap:bhh}: Bayesian Hyper-Heuristic}
\label{sec:symbols:bhh}

\begin{description}
	\item [\parbox{2cm}{$\alpha_{1}$}]
	\item [\parbox{2cm}{$\alpha_{k}$}]
	\item [\parbox{2cm}{$\alpha_{K}$}]
	\item [\parbox{2cm}{$\alpha$}]
	\item [\parbox{2cm}{$\beta_{1}$}]
	\item [\parbox{2cm}{$\beta_{2}$}]
	\item [\parbox{2cm}{$\beta_{j,k}$}]
	\item [\parbox{2cm}{$\beta_{K}^{J}$}]
	\item [\parbox{2cm}{$\beta$}]
	\item [\parbox{2cm}{$\boldsymbol{\alpha}$}]
	\item [\parbox{2cm}{$\boldsymbol{\beta}$}]
	\item [\parbox{2cm}{$\boldsymbol{\gamma}$}]
	\item [\parbox{2cm}{$\boldsymbol{\hat{\phi}}$}]
	\item [\parbox{2cm}{$\boldsymbol{\hat{\psi}}$}]
	\item [\parbox{2cm}{$\boldsymbol{\hat{\theta}}$}]
	\item [\parbox{2cm}{$\boldsymbol{\lambda}$}]
	\item [\parbox{2cm}{$\boldsymbol{\phi}$}]
	\item [\parbox{2cm}{$\boldsymbol{\psi}$}]
	\item [\parbox{2cm}{$\boldsymbol{\theta}$}]
	\item [\parbox{2cm}{$\boldsymbol{C}$}]
	\item [\parbox{2cm}{$\boldsymbol{E}$}]
	\item [\parbox{2cm}{$\boldsymbol{H}$}]
	\item [\parbox{2cm}{$\epsilon(-1)$}]
	\item [\parbox{2cm}{$\epsilon$}]
	\item [\parbox{2cm}{$\eta$}]
	\item [\parbox{2cm}{$\gamma_{0}$}]
	\item [\parbox{2cm}{$\gamma_{1,k}$}]
	\item [\parbox{2cm}{$\gamma_{1}$}]
	\item [\parbox{2cm}{$\gamma_{2,k}$}]
	\item [\parbox{2cm}{$\gamma_{K}$}]
	\item [\parbox{2cm}{$\hat{\phi}_{j,k}$}]
	\item [\parbox{2cm}{$\hat{\psi_{k}}$}]
	\item [\parbox{2cm}{$\hat{\psi}_{k}$}]
	\item [\parbox{2cm}{$\hat{\theta_{k}}$}]
	\item [\parbox{2cm}{$\hat{\theta}_{k}$}]
	\item [\parbox{2cm}{$\lambda_{1}$}]
	\item [\parbox{2cm}{$\lambda_{j}$}]
	\item [\parbox{2cm}{$\lambda_{J}$}]
	\item [\parbox{2cm}{$\mathbbm{1}_{0}$}]
	\item [\parbox{2cm}{$\mathbbm{1}_{1}$}]
	\item [\parbox{2cm}{$\mathbin{\%}$}]
	\item [\parbox{2cm}{$\mathcal{A}(v)$}]
	\item [\parbox{2cm}{$\mathcal{L}$}]
	\item [\parbox{2cm}{$\phi_{j,k}$}]
	\item [\parbox{2cm}{$\phi_{jk}$}]
	\item [\parbox{2cm}{$\phi$}]
	\item [\parbox{2cm}{$\psi_{k}$}]
	\item [\parbox{2cm}{$\psi_{k}N_{k}$}]
	\item [\parbox{2cm}{$\psi$}]
	\item [\parbox{2cm}{$\theta_{k}$}]
	\item [\parbox{2cm}{$\theta$}]
	\item [\parbox{2cm}{$A$}]
	\item [\parbox{2cm}{$B$}]
	\item [\parbox{2cm}{$Beta$}]
	\item [\parbox{2cm}{$Bin$}]
	\item [\parbox{2cm}{$c_{1}$}]
	\item [\parbox{2cm}{$c_{i,k}$}]
	\item [\parbox{2cm}{$c_{i}$}]
	\item [\parbox{2cm}{$Dir$}]
	\item [\parbox{2cm}{$e_{i,j,k}$}]
	\item [\parbox{2cm}{$e_{i}$}]
	\item [\parbox{2cm}{$e_{j}$}]
	\item [\parbox{2cm}{$E[\phi_{j,k}]$}]
	\item [\parbox{2cm}{$E[\psi_{k}]$}]
	\item [\parbox{2cm}{$E[\theta_{k}]$}]
	\item [\parbox{2cm}{$h_{i,k}$}]
	\item [\parbox{2cm}{$h_{i}$}]
	\item [\parbox{2cm}{$h_{k}$}]
	\item [\parbox{2cm}{$i$}]
	\item [\parbox{2cm}{$I$}]
	\item [\parbox{2cm}{$j$}]
	\item [\parbox{2cm}{$J$}]
	\item [\parbox{2cm}{$K-1$}]
	\item [\parbox{2cm}{$k$}]
	\item [\parbox{2cm}{$K$}]
	\item [\parbox{2cm}{$L$}]
	\item [\parbox{2cm}{$LSE$}]
	\item [\parbox{2cm}{$Mult$}]
	\item [\parbox{2cm}{$N_{0,k}$}]
	\item [\parbox{2cm}{$N_{1,k}$}]
	\item [\parbox{2cm}{$N_{j,k}$}]
	\item [\parbox{2cm}{$N_{j}$}]
	\item [\parbox{2cm}{$N_{k}$}]
	\item [\parbox{2cm}{$N$}]
	\item [\parbox{2cm}{$S$}]
	\item [\parbox{2cm}{$t$}]
\end{description}

\section{Chapter~\ref{chap:methodology}: Methodology}
\label{sec:symbols:methodology}

\begin{description}
	\item [\parbox{2cm}{$\alpha$}] Used as a scaling parameter for the \acs{LReLU} activation function. Also used as a threshold parameter for the statistical significance tests.
	\item [\parbox{2cm}{$\eta$}] Learning rate.
	\item [\parbox{2cm}{$\mathbb{R}$}] Real-number space.
	\item [\parbox{2cm}{$v_{ij}$}] The $j$-th dimension of the velocity vector of the $i$-th particle in the population implemented by \acs{PSO}.
	\item [\parbox{2cm}{$x_{ij}$}] The $j$-th dimension of the position vector of the $i$-th particle in the population implemented by \acs{PSO}.
\end{description}

\section{Chapter~\ref{chap:results}: Results}
\label{sec:symbols:results}

\begin{description}
	\item [\parbox{2cm}{$\alpha_{0}$}]
	\item [\parbox{2cm}{$\alpha_{6}$}]
	\item [\parbox{2cm}{$\alpha_{7}$}]
	\item [\parbox{2cm}{$\alpha_{8}$}]
	\item [\parbox{2cm}{$\alpha_{i} = 50$}]
	\item [\parbox{2cm}{$\boldsymbol{{H}}$}]
	\item [\parbox{2cm}{$\boldsymbol{\alpha}$}]
	\item [\parbox{2cm}{$\boldsymbol{\phi}$}]
	\item [\parbox{2cm}{$\boldsymbol{\psi}$}]
	\item [\parbox{2cm}{$\boldsymbol{\theta}$}]
	\item [\parbox{2cm}{$\boldsymbol{C}$}]
	\item [\parbox{2cm}{$\boldsymbol{E}$}]
	\item [\parbox{2cm}{$\boldsymbol{H}$}]
	\item [\parbox{2cm}{$\pm$}]
	\item [\parbox{2cm}{$\theta_{0}$}]
	\item [\parbox{2cm}{$\theta_{6}$}]
	\item [\parbox{2cm}{$\theta_{7}$}]
	\item [\parbox{2cm}{$\theta_{8}$}]
	\item [\parbox{2cm}{$c_{1}$}]
	\item [\parbox{2cm}{$Cat$}]
	\item [\parbox{2cm}{$Dir$}]
	\item [\parbox{2cm}{$e_{0}$}]
	\item [\parbox{2cm}{$h_{0}$}]
	\item [\parbox{2cm}{$h_{6}$}]
	\item [\parbox{2cm}{$h_{7}$}]
	\item [\parbox{2cm}{$h_{8}$}]
	\item [\parbox{2cm}{$i$}]
	\item [\parbox{2cm}{$K$}]
\end{description}