### 2023-01-11 article/sections/02-artificial-neural-networks.tex

\section{Artifical Neural Networks}
\label{sec:anns}