### 2023-02-01 article/appendices/c-datasets.tex
🚧 (article): Save point, WIP

\section{Datasets}
\label{app:appendix_datasets}

\begin{table}[H]
	\centering
	\caption{Classification datasets}
	\label{tab:methodology:datasets:classification}%
	\par\bigskip
	\resizebox{\textwidth}{!}{
		\begin{tabular}{ccccccccc}
			\textbf{dataset} & \textbf{output} & \textbf{types}             & \textbf{attributes} & \textbf{classes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation}          \\
			\midrule
			iris             & multivariate    & real                       & 4                   & 3                & 150                & 16             & 10             & ~\cite{ref:fisher:1936}    \\
			car              & multivariate    & categorical                & 6                   & 4                & 1728               & 128            & 14             & ~\cite{ref:bohanec:1988}   \\
			abalone          & multivariate    & categorical, integer, real & 8                   & 28               & 4177               & 256            & 17             & ~\cite{ref:waugh:1995}     \\
			wine quality     & multivariate    & real                       & 12                  & 11               & 4898               & 256            & 20             & ~\cite{ref:cortez:2009}    \\
			mushroom         & multivariate    & categorical                & 22                  & 2                & 8214               & 512            & 17             & ~\cite{ref:schlimmer:1987} \\
			bank             & multivariate    & real                       & 17                  & 2                & 45211              & 512            & 89             & ~\cite{ref:moro:2014}      \\
			diabetic         & multivariate    & integer                    & 55                  & 3                & 100000             & 1024           & 98             & ~\cite{ref:strack:2014}    \\
		\end{tabular}%
	}
\end{table}%

\begin{table}[htbp]
	\centering
	\caption{Regression datasets}
	\label{tab:methodology:datasets:regression}%
	\par\bigskip
	\resizebox{\textwidth}{!}{
		\begin{tabular}{cccccccc}
			\textbf{dataset}    & \textbf{output}           & \textbf{types} & \textbf{attributes} & \textbf{instances} & \textbf{batch} & \textbf{steps} & \textbf{citation}         \\
			\midrule
			fish toxicity       & multivariate              & real           & 7                   & 908                & 64             & 15             & ~\cite{ref:cassotti:2015} \\
			housing             & univariate                & real           & 13                  & 506                & 32             & 16             & ~\cite{ref:harrison:1978} \\
			forest fires        & multivariate              & real           & 13                  & 517                & 32             & 17             & ~\cite{ref:cortez:2007}   \\
			student performance & multivariate              & integer        & 33                  & 649                & 32             & 21             & ~\cite{ref:cortez:2008}   \\
			parkinsons          & multivariate              & integer, real  & 26                  & 5875               & 256            & 23             & ~\cite{ref:tsanas:2009}   \\
			air quality         & multivariate, time series & real           & 15                  & 9358               & 256            & 37             & ~\cite{ref:de:2008}       \\
			bike                & univariate                & integer, real  & 16                  & 17389              & 256            & 68             & ~\cite{ref:fanaee:2014}   \\
		\end{tabular}%
	}
\end{table}%