### 2023-02-04 article/main.tex

\documentclass[preprint,12pt]{elsarticle}
%% Use the option review to obtain double line spacing
% \documentclass[preprint,review,12pt]{elsarticle}

\input{journal/journal.tex}

\input{latex/packages.tex}
\input{latex/configuration.tex}
\input{acronyms/acronyms.tex}

\begin{document}

\input{frontmatter/frontmatter.tex}
\input{sections/01-introduction.tex}
\input{sections/02-artificial-neural-networks.tex}
\input{sections/03-heuristics.tex}
\input{sections/04-hyper-heuristics.tex}
\input{sections/05-probability.tex}
\input{sections/06-bayesian-hyper-heuristics.tex}
\input{sections/07-methodology.tex}
\input{sections/08-results.tex}
\input{sections/09-conclusion.tex}

\appendix

\input{appendices/c-datasets.tex}
\input{appendices/d-models.tex}
\input{appendices/e-heuristics.tex}
\input{appendices/f-proxies.tex}
\input{appendices/g-acronyms.tex}
\input{appendices/h-results-bhh-vs-heuristics-loss.tex}
\input{appendices/i-results-bhh-vs-heuristics-rank.tex}
\input{appendices/j-results-bhh-vs-heuristics-metrics.tex}

\input{references/bibliography.tex}

\end{document}

\endinput