### 2023-02-04 article/sections/03-heuristics.tex

\section{Heuristics}
\label{sec:heuristics}

A heuristic refers to an algorithmic search technique that serves as a guide to a search process where good solutions to a optimisation problem is being sought out. Many different techniques have been used to train \acp{FFNN} \cite{ref:kingma:2014}. At the time of writing, the majority of work that is published around the training of \acp{FFNN} involves the use of gradient-based techniques \cite{ref:nel:2021}.

Gradient-based heuristics are optimisation techniques that make use of derivates obtained from evaluating the \acs{ANN} being trained. In the context of supervised learning, loss functions produce a scalar value $\epsilon$, that represents the error between the output of the \acs{ANN} and the desired output. When using \acf{GD} to train \acp{ANN}, the loss function is used to adjust the weights of the \acs{ANN} in order to minimise the error \cite{ref:engelbrecht:2007}.

There are many variants of gradient-based heuristics, however, they all fundamentally apply the same generic \acs{GD} framework that propagates the error signal backwards through the \acs{ANN}. This algorithm is known as \acf{BP}. \acs{BP} was popularised by \citeauthor{ref:werbos:1994} \cite{ref:werbos:1994}.

The simplest type of \acs{GD} algorithm is referred to as \acs{SGD}, which implements a gradient-based weight update step for each training pattern. In the context of this research, the implementation of \acs{SGD} refers to the mini-batch training implementation of \acs{GD}, where a small batch of training patterns are fed to the \acs{FFNN} at once and the error function is aggregated across all training patterns.

Alternative variants have been proposed that lead to better control over the convergence characteristics caused by \acs{SGD}. This research focusses on a number of these variants that include \acf{Momentum} \cite{ref:qian:1999}, \acf{NAG} \cite{ref:sutskever:2013}, \acf{Adagrad} \cite{ref:duchi:2011}, \acs{Adadelta} \cite{ref:zeiler:2012}, \acf{RMSProp} \cite{ref:hinton:2012} and \acf{Adam} \cite{ref:kingma:2014}.

Gradient-based heuristics are sensitive to the problem that they are applied to, with hyper-parameter selection often dominating the research focus \cite{ref:bengio:2000, ref:feurer:2019}. \citeauthor{ref:blum:2003} \cite{ref:blum:2003} mention that since the 1980's, a new kind of approximate algorithm has emerged which tries to combine basic heuristic methods in higher level frameworks aimed at efficiently and effectively exploring a search space. These methods are referred to as \acp{MH}.

The biggest difference between \acp{MH} and gradient-bases heuristics is that
\acp{MH} make use of meta-information obtained as a result of evaluating the \acs{FFNN} during training and is not limited to information about the search space \cite{ref:blum:2003}. Similar to gradient-based heuristics, a number of different meta-heuristics have been used to successfully train \acp{FFNN} \cite{ref:rakitianskaia:2012, ref:vanwyk:2014, ref:espinal:2011, ref:gupta:1999}. This research takes a particular interest in population-based \acp{MH} that have been used to successfully train \acp{FFNN}. These include \acf{PSO} \cite{ref:shi:1998}, \acf{DE} \cite{ref:price:2006} and \acfp{GA} \cite{ref:fraser:1957}.