### 2023-02-05 article/frontmatter/abstract.tex
✨ (article): Done, submitting for review

\begin{abstract}
	This research introduces a novel population-based \acf{BHH} that is used to train \acfp{FFNN}. An in-depth behaviour analysis is done and the performance of the \acs{BHH} is compared to that of ten popular low-level heuristics, each with different search behaviours. The chosen heuristic pool consists of classic gradient-based heuristics as well as \acf{MH}. The empirical process is executed on fourteen datasets consisting of classification and regression problems with varying characteristics. The \acs{BHH} is shown to be able to train \acp{FFNN} well and provide an automated method for finding the best heuristic to train the \acp{FFNN} at various stages of the training process.
\end{abstract}