### 2025-01-13 dissertation/appendix/datasets.tex
💚 For Taylah, Beané & Aleah

\chapter{Datasets}
\label{app:datasets}

This appendix provides insight into the preparation and usage of datasets that were used as part of the empirical process. The data processing tasks included:

\begin{itemize}
	\item Exploratory analysis and visualisation off data distribution.

	\item Dropping of rows that contain missing data in and standardisation of data types.

	\item Normalising all numerical features using the \index{standard score scaler}standard score scaler and one-hot encoding of all categorical features.

	\item Splitting the datasets into a training set (80\%) and a testing set (20\%) and then shuffling and batching of data.
\end{itemize}