### 2025-01-13 dissertation/appendix/symbols.tex

\chapter{Symbols}
\label{app:symbols}

This appendix summarises the important symbols used throughout the dissertation. A short description is provided for each symbol. Symbols are divided according to the chapter in which they were introduced and are ordered alphabetically.

\section{Chapter~\ref{chap:anns}: Artificial Neural Networks}
\label{sec:symbols:anns}

\begin{description}
	% \item
	%       [\parbox{2cm}{$EXAMPLE$}] \parbox{12.5cm}{THIS IS AN EXAMPLE DESCRIPTION. \lbrack Eq.~\eqref{eq:an_function_mapping} pg.~\pageref{eq:an_function_mapping} \rbrack}

	\item [\parbox{2cm}{$\alpha$}] \parbox{12.5cm}{Scaling parameter used by the \acs{LReLU} activation function.}
	\item [\parbox{2cm}{$\beta$}] \parbox{12.5cm}{Mini-batch size.}
	\item [\parbox{2cm}{$\boldsymbol{v}$}] \parbox{12.5cm}{The weight vector.}
	\item [\parbox{2cm}{$\boldsymbol{x}_p$}] \parbox{12.5cm}{The $p$-th input vector/pattern.}
	\item [\parbox{2cm}{$\boldsymbol{x}$}] \parbox{12.5cm}{The input pattern/vector.}
	\item [\parbox{2cm}{$\boldsymbol{y}_{p}$}] \parbox{12.5cm}{The output vector for pattern $p$.}
	\item [\parbox{2cm}{$\boldsymbol{y}$}] \parbox{12.5cm}{The output vector.}
	\item [\parbox{2cm}{$\boldsymbol{y^{'}}$}] \parbox{12.5cm}{The softmax form of the output vector.}
	\item [\parbox{2cm}{$\epsilon$}] \parbox{12.5cm}{The cost/loss produced by the loss function.}
	\item [\parbox{2cm}{$\hat{y}_{k,p}$}] \parbox{12.5cm}{The target output for the \acs{ANN} at the $k$-th dimension for the $p$-th pattern.}
	\item [\parbox{2cm}{$\lambda$}] \parbox{12.5cm}{Scaling/control parameter used by the sigmoid and tanh activation functions.}
	\item [\parbox{2cm}{$\mathbb{R}^{I}$}] \parbox{12.5cm}{The real-number space in I dimensions.}
	\item [\parbox{2cm}{$\mathbb{R}^{K}$}] \parbox{12.5cm}{The real-number space in $K$ dimensions.}
	\item [\parbox{2cm}{$\mathbb{R}^{T}$}] \parbox{12.5cm}{The real-number space in T dimensions.}
	\item [\parbox{2cm}{$\mathbb{R}$}] \parbox{12.5cm}{The real-number space.}
	\item [\parbox{2cm}{$\mathbbm{1}$}] \parbox{12.5cm}{The indicator function.}
	\item [\parbox{2cm}{$\mu_i$}] \parbox{12.5cm}{The mean of the input at dimension $i$.}
	\item [\parbox{2cm}{$\omega_{i}$}] \parbox{12.5cm}{The general weight vector used for normalisation at dimension $i$.}
	\item [\parbox{2cm}{$\omega_{max}$}] \parbox{12.5cm}{The upper bound of the uniform distribution used for weights initialisation.}
	\item [\parbox{2cm}{$\omega_{min}$}] \parbox{12.5cm}{The lower bound of the uniform distribution used for weights initialisation.}
	\item [\parbox{2cm}{$\omega$}] \parbox{12.5cm}{The general weight vector used for normalisation.}
	\item [\parbox{2cm}{$\sigma^2_i$}] \parbox{12.5cm}{The unit variance of the input at dimension $i$.}
	\item [\parbox{2cm}{$\sigma$}] \parbox{12.5cm}{The standard deviation of the truncated normal distribution used for weight initialisation.}
	\item [\parbox{2cm}{$\tau$}] \parbox{12.5cm}{Phase shift threshold parameter.}
	\item [\parbox{2cm}{$\theta$}] \parbox{12.5cm}{The bias unit.}
	\item [\parbox{2cm}{$c$}] \parbox{12.5cm}{Index used for $c$-th class.}
	\item [\parbox{2cm}{$C$}] \parbox{12.5cm}{Total number of classes.}
	\item [\parbox{2cm}{$f_{AN}$}] \parbox{12.5cm}{The mapping function realised by the \acs{AN}.}
	\item [\parbox{2cm}{$f(net)$}] \parbox{12.5cm}{The activation function over net input signal.}
	\item [\parbox{2cm}{$f(x)$}] \parbox{12.5cm}{The function used to describe the \acs{LReLU} activation function.}
	\item [\parbox{2cm}{$f$}] \parbox{12.5cm}{A shortened form of the activation function.}
	\item [\parbox{2cm}{$fanin$}] \parbox{12.5cm}{The number of input neurons to the weight vector.}
	\item [\parbox{2cm}{$fanout$}] \parbox{12.5cm}{The number of output neurons from the weight vector.}
	\item [\parbox{2cm}{$h_{j}$}] \parbox{12.5cm}{The $j$-th dimension of the hidden layer.}
	\item [\parbox{2cm}{$i$}] \parbox{12.5cm}{Index for the input vector.}
	\item [\parbox{2cm}{$I$}] \parbox{12.5cm}{The input pattern's dimension size.}
	\item [\parbox{2cm}{$j$}] \parbox{12.5cm}{Index used for the hidden layer.}
	\item [\parbox{2cm}{$k$}] \parbox{12.5cm}{Index for the output vector.}
	\item [\parbox{2cm}{$K$}] \parbox{12.5cm}{Number of output units.}
	\item [\parbox{2cm}{$net_{h,y}$}] \parbox{12.5cm}{The net input signal at the output layer.}
	\item [\parbox{2cm}{$net_{i,h}$}] \parbox{12.5cm}{The net input signal at the hidden layer.}
	\item [\parbox{2cm}{$net^{'}$}] \parbox{12.5cm}{The augmented net input signal that includes the bias term.}
	\item [\parbox{2cm}{$net$}] \parbox{12.5cm}{The net input signal to the \acs{AN}.}
	\item [\parbox{2cm}{$p$}] \parbox{12.5cm}{Index for the input data patterns.}
	\item [\parbox{2cm}{$P$}] \parbox{12.5cm}{Total number of input data patterns.}
	\item [\parbox{2cm}{$T$}] \parbox{12.5cm}{The target pattern's dimension size.}
	\item [\parbox{2cm}{$v_{i,j}$}] \parbox{12.5cm}{The weight associated with input node $x_{i}$ and the hidden node $h_{j}$.}
	\item [\parbox{2cm}{$v_{i}$}] \parbox{12.5cm}{The $i$-th dimension of the weight vector.}
	\item [\parbox{2cm}{$v_{i+1}$}] \parbox{12.5cm}{The $i+1$-th dimension of the weight vector.}
	\item [\parbox{2cm}{$w_{j,k}$}] \parbox{12.5cm}{The weight associated with hidden node $h_{j}$ and the output node $y_{k}$.}
	\item [\parbox{2cm}{$x_{i_{max}}$}] \parbox{12.5cm}{The input's maximum value at dimension $i$.}
	\item [\parbox{2cm}{$x_{i_{min}}$}] \parbox{12.5cm}{The input's minimum value at dimension $i$.}
	\item [\parbox{2cm}{$x_{i,p}^{'}$}] \parbox{12.5cm}{The normalised form of the input at dimension $i$, pattern $p$.}
	\item [\parbox{2cm}{$x_{i,p}$}] \parbox{12.5cm}{The original input at dimension $i$, pattern $p$.}
	\item [\parbox{2cm}{$x_{i}$}] \parbox{12.5cm}{The $i$-th dimension of the input pattern/vector.}
	\item [\parbox{2cm}{$x_{i+1}$}] \parbox{12.5cm}{The $i+1$-th dimension of the input pattern/vector.}
	\item [\parbox{2cm}{$x_{max}$}] \parbox{12.5cm}{The input's maximum value used by the min-max scaler.}
	\item [\parbox{2cm}{$x_{min}$}] \parbox{12.5cm}{The input's minimum value used by the min-max scaler.}
	\item [\parbox{2cm}{$x^{+}$}] \parbox{12.5cm}{The positive part of the input parameter's domain.}
	\item [\parbox{2cm}{$y_{k,p}$}] \parbox{12.5cm}{The output produced by the \acs{ANN} at the $k$-th dimension for the $p$-th pattern.}
	\item [\parbox{2cm}{$y_{k}$}] \parbox{12.5cm}{The $k$-th dimension of the output vector.}
	\item [\parbox{2cm}{$y^{'}_k$}] \parbox{12.5cm}{The softmax value of the output at dimension $k$.}

\end{description}

\section{Chapter~\ref{chap:heuristics}: Heuristics}
\label{sec:symbols:heuristics}

\begin{description}
	\item [\parbox{2cm}{$\alpha$}] \parbox{12.5cm}{The momentum hyper-parameter.}
	\item [\parbox{2cm}{$\beta_{1}$}] \parbox{12.5cm}{First decay rate parameter for \acs{Adam}.}
	\item [\parbox{2cm}{$\beta_{2}$}] \parbox{12.5cm}{Second decay rate parameter for \acs{Adam}.}
	\item [\parbox{2cm}{$\beta^{t}_{1}$}] \parbox{12.5cm}{First decay rate parameter for \acs{Adam} at time step $t$.}
	\item [\parbox{2cm}{$\beta^{t}_{2}$}] \parbox{12.5cm}{Second decay rate parameter for \acs{Adam} at time step $t$.}
	\item [\parbox{2cm}{$\beta$}] \parbox{12.5cm}{Scale factor as implemented by \acs{DE}.}
	\item [\parbox{2cm}{$\boldsymbol{\hat{y}}_{i}(t)$}] \parbox{12.5cm}{The $i$-th particle's global best position vector as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$\boldsymbol{\hat{y}}$}] \parbox{12.5cm}{Short form of the global best position vector.}
	\item [\parbox{2cm}{$\boldsymbol{G_{t}}$}] \parbox{12.5cm}{The sum of the squares of the gradients.}
	\item [\parbox{2cm}{$\boldsymbol{g}_{t}$}] \parbox{12.5cm}{The gradient vector at time step $t$.}
	\item [\parbox{2cm}{$\boldsymbol{g}^{2}_{t}$}] \parbox{12.5cm}{The squared gradient vector at time step $t$.}
	\item [\parbox{2cm}{$\boldsymbol{g}$}] \parbox{12.5cm}{General symbol used to represent the gradient vector, retrieved from the error function relative to the weight vector.}
	\item [\parbox{2cm}{$\boldsymbol{m}_{t}$}] \parbox{12.5cm}{The estimate of the first moment at time step $t$.}
	\item [\parbox{2cm}{$\boldsymbol{m}_{t+1}$}] \parbox{12.5cm}{The estimate of the first moment at time step $t+1$.}
	\item [\parbox{2cm}{$\boldsymbol{r}_{1}$}] \parbox{12.5cm}{The vector for the first stochastic scaling parameter as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$\boldsymbol{r}_{2}$}] \parbox{12.5cm}{The vector for the second stochastic scaling parameter as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$\boldsymbol{u}_{i}(t)$}] \parbox{12.5cm}{The $i$-th entity's trial vector at time step $t$ implemented by \acs{DE}.}
	\item [\parbox{2cm}{$\boldsymbol{v}_{i}(t)$}] \parbox{12.5cm}{The $i$-th particle's velocity vector at time step $t$ as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$\boldsymbol{V}_{max}$}] \parbox{12.5cm}{The upper bound of the velocity clamping vector.}
	\item [\parbox{2cm}{$\boldsymbol{v}_{t}$}] \parbox{12.5cm}{The estimate of the second moment at time step $t$.}
	\item [\parbox{2cm}{$\boldsymbol{v}_{t+1}$}] \parbox{12.5cm}{The estimate of the second moment at time step $t+1$.}
	\item [\parbox{2cm}{$\boldsymbol{v}$}] \parbox{12.5cm}{The velocity vector resulting from \acs{Momentum}.}
	\item [\parbox{2cm}{$\boldsymbol{w}_{t}^{2}$}] \parbox{12.5cm}{The squared weight vector at time step $t$.}
	\item [\parbox{2cm}{$\boldsymbol{w}_{t+1}$}] \parbox{12.5cm}{The weight vector at time step $t+1$.}
	\item [\parbox{2cm}{$\boldsymbol{w}$}] \parbox{12.5cm}{General symbol used to represent the weight vector.}
	\item [\parbox{2cm}{$\boldsymbol{w}$}] \parbox{12.5cm}{The weight vector.}
	\item [\parbox{2cm}{$\boldsymbol{x}_{i_{1}}(t)$}] \parbox{12.5cm}{The $i$-th entity's target vector at time step $t$ implemented by \acs{DE}.}
	\item [\parbox{2cm}{$\boldsymbol{x}_{i_{2}}(t)$}] \parbox{12.5cm}{The $i$-th entity's first selected individual's position vector at time step $t$ implemented by \acs{DE}.}
	\item [\parbox{2cm}{$\boldsymbol{x}_{i_{3}}(t)$}] \parbox{12.5cm}{The $i$-th entity's second selected individual's position vector at time step $t$ implemented by \acs{DE}.}
	\item [\parbox{2cm}{$\boldsymbol{x}_{i}(0)$}] \parbox{12.5cm}{The $i$-th particle's position vector at time step $0$ as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$\boldsymbol{x}_{i}(t)$}] \parbox{12.5cm}{The $i$-th particle's position vector at time step $t$ as implemented by \acs{PSO}. Also represents the $i$-th parent's position vector at time step $t$ as implemented by \acs{DE}.}
	\item [\parbox{2cm}{$\boldsymbol{x}_{i}$}] \parbox{12.5cm}{The $i$-th particle's position vector as implemented by \acs{PSO}. Also represents the $i$-th parent's position vector as implemented by \acs{DE}.}
	\item [\parbox{2cm}{$\boldsymbol{x}_{i}$}] \parbox{12.5cm}{The $i$-th particle's position vector as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$\boldsymbol{x}_{max}$}] \parbox{12.5cm}{The upper bound constraint vector of the position vector's values.}
	\item [\parbox{2cm}{$\boldsymbol{x}_{min}$}] \parbox{12.5cm}{The lower bound constraint vector of the position vector's values.}
	\item [\parbox{2cm}{$\boldsymbol{x}'_{i}(t)$}] \parbox{12.5cm}{The $i$-th offspring's position vector at time step $t$ as implemented by \acs{DE}.}
	\item [\parbox{2cm}{$\boldsymbol{y}_{i}(0)$}] \parbox{12.5cm}{The $i$-th particle's personal best position vector at time step $0$ as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$\boldsymbol{y}_{i}(t)$}] \parbox{12.5cm}{The $i$-th particle's personal best position vector at time step $t$ as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$\Delta \boldsymbol{w}^{2}_{t}$}] \parbox{12.5cm}{The delta of the squared weight vector at time step $t$.}
	\item [\parbox{2cm}{$\Delta w_{i}(t)$}] \parbox{12.5cm}{The change in the $i$-th dimension of the weight vector at time step $t$.}
	\item [\parbox{2cm}{$\emptyset$}] \parbox{12.5cm}{The empty set.}
	\item [\parbox{2cm}{$\epsilon_{T_{p}}$}] \parbox{12.5cm}{The error produced for pattern $p$ at time step $T$.}
	\item [\parbox{2cm}{$\epsilon_{T}$}] \parbox{12.5cm}{The error at time step $T$.}
	\item [\parbox{2cm}{$\epsilon$}] \parbox{12.5cm}{The error between the output of the \acs{FFNN} and the target pattern, produced by the loss function. Also used as a small error value used by heuristics to avoid division by 0.}
	\item [\parbox{2cm}{$\eta$}] \parbox{12.5cm}{Learning rate.}
	\item [\parbox{2cm}{$\hat{\boldsymbol{m}}_{t}$}] \parbox{12.5cm}{The bias-corrected first moment at time step $t$.}
	\item [\parbox{2cm}{$\hat{\boldsymbol{v}}_{t}$}] \parbox{12.5cm}{The bias-corrected second moment at time step $t$.}
	\item [\parbox{2cm}{$\hat{y}_{ij}(t)$}] \parbox{12.5cm}{The $i$-th particle's $j$-th dimension of the global best position vector at time step $t$.}
	\item [\parbox{2cm}{$\hat{y}_{ij}$}] \parbox{12.5cm}{The $i$-th particle's $j$-th dimension of the global best position vector.}
	\item [\parbox{2cm}{$\mathcal{C}(0)$}] \parbox{12.5cm}{The set of entities in the population at time step $0$.}
	\item [\parbox{2cm}{$\mathcal{C}(t)$}] \parbox{12.5cm}{The set of entities in the population at time step $t$.}
	\item [\parbox{2cm}{$\mathcal{C}(t+1)$}] \parbox{12.5cm}{The set of entities in the population at time step $t+1$.}
	\item [\parbox{2cm}{$\mathcal{C}$}] \parbox{12.5cm}{The set of entities in the population.}
	\item [\parbox{2cm}{$\mathcal{E}$}] \parbox{12.5cm}{The error function.}
	\item [\parbox{2cm}{$\mathcal{J}$}] \parbox{12.5cm}{The set of crossover points.}
	\item [\parbox{2cm}{$\odot$}] \parbox{12.5cm}{The matrix-vector product.}
	\item [\parbox{2cm}{$c_{1}$}] \parbox{12.5cm}{The cognitive control parameter as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$c_{2}$}] \parbox{12.5cm}{The social control parameter as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$E[\boldsymbol{g}^{2}]_{t - 1}$}] \parbox{12.5cm}{The expected value of the squared gradients at time step $t-1$.}
	\item [\parbox{2cm}{$E[\boldsymbol{g}^{2}]_{t}$}] \parbox{12.5cm}{The expected value of the squared gradients at time step $t$.}
	\item [\parbox{2cm}{$E[\Delta \boldsymbol{w}^{2}]_{t - 1}$}] \parbox{12.5cm}{The expected value of the delta of the squared weight vector at time step $t-1$.}
	\item [\parbox{2cm}{$E[\Delta \boldsymbol{w}^{2}]_{t}$}] \parbox{12.5cm}{The expected value of the delta of the squared weight vector at time step $t$.}
	\item [\parbox{2cm}{$f(\boldsymbol{\hat{y}})$}] \parbox{12.5cm}{The evaluation of the global best position.}
	\item [\parbox{2cm}{$f(\boldsymbol{x}_{i}(t))$}] \parbox{12.5cm}{The evaluation of the $i$-th entity's candidate solution at time step $t$.}
	\item [\parbox{2cm}{$f(\boldsymbol{x}'_{i}(t))$}] \parbox{12.5cm}{The evaluation of the $i$-th offspring's candidate solution at time step $t$.}
	\item [\parbox{2cm}{$f(\boldsymbol{y}_{i})$}] \parbox{12.5cm}{The evaluation of the $i$-th particle's personal best position.}
	\item [\parbox{2cm}{$f(\hat{y})$}] \parbox{12.5cm}{The evaluation of the global best position.}
	\item [\parbox{2cm}{$f(x_{i})$}] \parbox{12.5cm}{The evaluation of the $i$-th particle's solution with respect to the objective function.}
	\item [\parbox{2cm}{$f(x)$}] \parbox{12.5cm}{The objective function over $x$.}
	\item [\parbox{2cm}{$f$}] \parbox{12.5cm}{A shortened form of the objective function being optimised.}
	\item [\parbox{2cm}{$g_{t,i}$}] \parbox{12.5cm}{The $i$-th dimension of the gradient vector at time step $t$.}
	\item [\parbox{2cm}{$G_{t,i}$}] \parbox{12.5cm}{The sum of the squared gradients with regards to the $i$-th dimension of the weight vector.}
	\item [\parbox{2cm}{$G_{t,ii}$}] \parbox{12.5cm}{Diagonal matrix containing the sum of the squared gradients with regards to the $i$-th dimension of the weight vector.}
	\item [\parbox{2cm}{$i_{1}$}] \parbox{12.5cm}{Index for the target vector in \acs{DE}.}
	\item [\parbox{2cm}{$i_{2}$}] \parbox{12.5cm}{Index for the first selected individual in \acs{DE}.}
	\item [\parbox{2cm}{$i_{3}$}] \parbox{12.5cm}{Index for the second selected individual in \acs{DE}.}
	\item [\parbox{2cm}{$i$}] \parbox{12.5cm}{Index used for the weight vector. Also used as the index for the $i$-th particle in the population, implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$I$}] \parbox{12.5cm}{The total number of particles in the swarm (particle swarm size).}
	\item [\parbox{2cm}{$j^{*}$}] \parbox{12.5cm}{Randomly selected crossover point.}
	\item [\parbox{2cm}{$j$}] \parbox{12.5cm}{Index used for the dimensions of candidate solutions.}
	\item [\parbox{2cm}{$J$}] \parbox{12.5cm}{The total number of dimensions in the position vector for particles in \acs{PSO}.}
	\item [\parbox{2cm}{$k$}] \parbox{12.5cm}{Index used for the output layer.}
	\item [\parbox{2cm}{$K$}] \parbox{12.5cm}{Total number of output units.}
	\item [\parbox{2cm}{$m_{j}(t)$}] \parbox{12.5cm}{The $j$-th dimension of the crossover mask, $m$, at time step $t$.}
	\item [\parbox{2cm}{$N_{t}$}] \parbox{12.5cm}{Tournament selection size.}
	\item [\parbox{2cm}{$N$}] \parbox{12.5cm}{An alternative to the particle swarm size.}
	\item [\parbox{2cm}{$net_{p}$}] \parbox{12.5cm}{The net input signal for pattern $p$.}
	\item [\parbox{2cm}{$o_{i,p}$}] \parbox{12.5cm}{The $i$-th dimension of the output for pattern $p$.}
	\item [\parbox{2cm}{$o_{k,p}$}] \parbox{12.5cm}{The $k$-th dimension of the output for pattern $p$.}
	\item [\parbox{2cm}{$p_{c}$}] \parbox{12.5cm}{Probability of producing offspring.}
	\item [\parbox{2cm}{$p_{m}$}] \parbox{12.5cm}{The mutation probability.}
	\item [\parbox{2cm}{$p_{r}$}] \parbox{12.5cm}{The crossover probability i.e. recombination probability.}
	\item [\parbox{2cm}{$p_{x}$}] \parbox{12.5cm}{Bitswapping probability.}
	\item [\parbox{2cm}{$p$}] \parbox{12.5cm}{Index used for input data patterns.}
	\item [\parbox{2cm}{$r_{1_{j}}(t)$}] \parbox{12.5cm}{The $j$-th dimension of the first stochastic scaling parameter at time step $t$ as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$r_{2_{j}}(t)$}] \parbox{12.5cm}{The $j$-th dimension of the second stochastic scaling parameter at time step $t$ as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$RMS[\boldsymbol{g}]_{t}$}] \parbox{12.5cm}{The root mean squared gradient vector at time step $t$.}
	\item [\parbox{2cm}{$RMS$}] \parbox{12.5cm}{The root mean squared error criterion.}
	\item [\parbox{2cm}{$t_{i,p}$}] \parbox{12.5cm}{The $i$-th dimension of the target for pattern $p$.}
	\item [\parbox{2cm}{$t_{k,p}$}] \parbox{12.5cm}{The $k$-th dimension of the target for pattern $p$.}
	\item [\parbox{2cm}{$t$}] \parbox{12.5cm}{Time step.}
	\item [\parbox{2cm}{$T$}] \parbox{12.5cm}{Time step.}
	\item [\parbox{2cm}{$u_{ij}(t)$}] \parbox{12.5cm}{The $j$-th dimension of the $i$-th entity's trial vector at time step $t$ implemented by \acs{DE}.}
	\item [\parbox{2cm}{$U$}] \parbox{12.5cm}{The uniform distribution.}
	\item [\parbox{2cm}{$v_{ij}(t)$}] \parbox{12.5cm}{The $i$-th particle's $j$-th dimension of the velocity vector at time step $t$. Also used in the mutation operator for \acp{GA} to represent some sampled mutation update value.}
	\item [\parbox{2cm}{$v_{ij}(t+1)$}] \parbox{12.5cm}{The $i$-th particle's $j$-th dimension of the velocity vector at time step $t+1$.}
	\item [\parbox{2cm}{$V_{max,j}$}] \parbox{12.5cm}{The $j$-th dimension of the upper bound of the velocity clamping vector.}
	\item [\parbox{2cm}{$v'_{ij}(t+1)$}] \parbox{12.5cm}{The $i$-th particle's $j$-th dimension of the clamped velocity vector at time step $t+1$.}
	\item [\parbox{2cm}{$w_{i}(t-1)$}] \parbox{12.5cm}{The $i$-th dimension of the weight vector at time step $t-1$.}
	\item [\parbox{2cm}{$w_{i}(t)$}] \parbox{12.5cm}{The $i$-th dimension of the weight vector at time step $t$.}
	\item [\parbox{2cm}{$w_{i}$}] \parbox{12.5cm}{The $i$-th dimension of the weight vector.}
	\item [\parbox{2cm}{$w_{t,i}$}] \parbox{12.5cm}{The $i$-th dimension of the weight vector at time step $t$.}
	\item [\parbox{2cm}{$w_{t+1,i}$}] \parbox{12.5cm}{The $i$-th dimension of the weight vector at time step $t+1$.}
	\item [\parbox{2cm}{$w$}] \parbox{12.5cm}{The inertia weight hyper-parameter as implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$W$}] \parbox{12.5cm}{Window size of previous squared gradients.}
	\item [\parbox{2cm}{$x_{ij}(t)$}] \parbox{12.5cm}{The $i$-th particle's/entity's $j$-th dimension of the position vector at time step $t$.}
	\item [\parbox{2cm}{$x_{ij}(t+1)$}] \parbox{12.5cm}{The $i$-th particle's $j$-th dimension of the position vector at time step $t+1$.}
	\item [\parbox{2cm}{$x_{ij}$}] \parbox{12.5cm}{The $i$-th particle's/entity's $j$-th dimension of the position vector.}
	\item [\parbox{2cm}{$x_{max,j}$}] \parbox{12.5cm}{The $j$-th dimension of the upper bound constraint for an entity's position.}
	\item [\parbox{2cm}{$x_{min,j}$}] \parbox{12.5cm}{The $j$-th dimension of the lower bound constraint for an entity's position.}
	\item [\parbox{2cm}{$x'_{ij}(t)$}] \parbox{12.5cm}{The $i$-th offspring's $j$-th dimension of the position vector at time step $t$.}
	\item [\parbox{2cm}{$x$}] \parbox{12.5cm}{The independent variables to the objective function. Also used in \acs{DE} notation to denote the selection mechanism for the trial vector.}
	\item [\parbox{2cm}{$y_{ij}(t)$}] \parbox{12.5cm}{The $i$-th particle's $j$-th dimension of the personal best position vector at time step $t$.}
	\item [\parbox{2cm}{$y_{ij}$}] \parbox{12.5cm}{The $i$-th particle's $j$-th dimension of the personal best position vector.}
	\item [\parbox{2cm}{$y$}] \parbox{12.5cm}{Used in \acs{DE} notation to denote the number of difference vectors to include.}
	\item [\parbox{2cm}{$z_{i,p}$}] \parbox{12.5cm}{The input value at index $i$ for pattern $p$.}
	\item [\parbox{2cm}{$z$}] \parbox{12.5cm}{Used in \acs{DE} notation to denote the type of crossover operator to use.}
\end{description}

\section{Chapter~\ref{chap:hhs}: Hyper-Heuristics}
\label{sec:symbols:hhs}

\begin{description}
	\item [\parbox{2cm}{$K$}] \parbox{12.5cm}{The number of folds to execute for $K$-fold crossover validation.}
\end{description}

\section{Chapter~\ref{chap:probability}: Probability}
\label{sec:symbols:probability}

\begin{description}
	\item [\parbox{2cm}{$\alpha_{0}$}] \parbox{12.5cm}{The sum of all the values in each dimension of the concentration parameter $\boldsymbol{\alpha}$.}
	\item [\parbox{2cm}{$\alpha_{1}$}] \parbox{12.5cm}{The first dimension of the concentration parameter $\boldsymbol{\alpha}$.}
	\item [\parbox{2cm}{$\alpha_{k}$}] \parbox{12.5cm}{The $k$-th dimension of the concentration parameter $\boldsymbol{\alpha}$.}
	\item [\parbox{2cm}{$\alpha_{K}$}] \parbox{12.5cm}{The last dimension of the concentration parameter $\boldsymbol{\alpha}$.}
	\item [\parbox{2cm}{$\alpha'$}] \parbox{12.5cm}{The update form of the prior parameter, $\alpha$, for a Beta probability distribution.}
	\item [\parbox{2cm}{$\alpha$}] \parbox{12.5cm}{Shape parameter for the Beta probability distribution. Also used for confidence levels for the statistical tests.}
	\item [\parbox{2cm}{$\bar{A}$}] \parbox{12.5cm}{Not observing random event $A$.}
	\item [\parbox{2cm}{$\beta'$}] \parbox{12.5cm}{The update form of the prior parameter, $\beta$, for a Beta probability distribution.}
	\item [\parbox{2cm}{$\beta$}] \parbox{12.5cm}{Shape parameter for the Beta probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\alpha'}$}] \parbox{12.5cm}{Update form of the prior parameter, $\boldsymbol{\alpha}$, for a Dirichlet probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\alpha}$}] \parbox{12.5cm}{The concentration parameter vector that parameterises the Dirichlet probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\mathcal{D}}$}] \parbox{12.5cm}{All the prior data of $\boldsymbol{X}$}
	\item [\parbox{2cm}{$\boldsymbol{\theta}$}] \parbox{12.5cm}{The probability vector in multiple dimensions used by multiple probability distributions.}
	\item [\parbox{2cm}{$\boldsymbol{N}$}] \parbox{12.5cm}{The counts for each occurrence of a category $k$ over $I$ independent, identical (iid) events.}
	\item [\parbox{2cm}{$\boldsymbol{x_{1}}$}] \parbox{12.5cm}{The outcomes of the first random event in $I$ independent, identical (iid) events for $k$ categories.}
	\item [\parbox{2cm}{$\boldsymbol{x_{i}}$}] \parbox{12.5cm}{The outcomes of the $i$-th random event in $I$ independent, identical (iid) events for $k$ categories.}
	\item [\parbox{2cm}{$\boldsymbol{x_{I}}$}] \parbox{12.5cm}{The outcomes of the last random event in $I$ independent, identical (iid) events for $k$ categories.}
	\item [\parbox{2cm}{$\boldsymbol{X}$}] \parbox{12.5cm}{Matrix of Bernoulli probability distributions. Also used to represent the outcomes of $I$ independent, identical (iid) random events.}
	\item [\parbox{2cm}{$\boldsymbol{x}$}] \parbox{12.5cm}{The random variable for various probability distributions in multiple dimensions.}
	\item [\parbox{2cm}{$\boldsymbol{Y}$}] \parbox{12.5cm}{A vector of random events.}
	\item [\parbox{2cm}{$\emptyset$}] \parbox{12.5cm}{The empty set.}
	\item [\parbox{2cm}{$\Gamma'(n)$}] \parbox{12.5cm}{The first derivative of the Gamma function over input $n$.}
	\item [\parbox{2cm}{$\Gamma(n)$}] \parbox{12.5cm}{The Gamma function over input $n$.}
	\item [\parbox{2cm}{$\Gamma$}] \parbox{12.5cm}{The Gamma function.}
	\item [\parbox{2cm}{$\mathbb{R}$}] \parbox{12.5cm}{Real-number space.}
	\item [\parbox{2cm}{$\mathbb{R}$}] \parbox{12.5cm}{The real-number space.}
	\item [\parbox{2cm}{$\mathbbm{1}$}] \parbox{12.5cm}{Indicator function.}
	\item [\parbox{2cm}{$\mathcal{A}(v)$}] \parbox{12.5cm}{Functional form of priors.}
	\item [\parbox{2cm}{$\mathcal{L}$}] \parbox{12.5cm}{The likelihood function.}
	\item [\parbox{2cm}{$\overset{\text{iid}}{\sim}$}] \parbox{12.5cm}{Independent, identical (iid) random events.}
	\item [\parbox{2cm}{$\psi$}] \parbox{12.5cm}{The logarithmic derivative of the Gamma function, called the Digamma Function.}
	\item [\parbox{2cm}{$\theta_{1}$}] \parbox{12.5cm}{The first dimension of the probability vector $\boldsymbol{\theta}$.}
	\item [\parbox{2cm}{$\theta_{k}$}] \parbox{12.5cm}{The $k$-th dimension of the probability vector $\boldsymbol{\theta}$.}
	\item [\parbox{2cm}{$\theta_{K}$}] \parbox{12.5cm}{The last dimension of the probability vector $\boldsymbol{\theta}$.}
	\item [\parbox{2cm}{$\theta$}] \parbox{12.5cm}{Success probability for the Bernoulli probability distribution.}
	\item [\parbox{2cm}{$A$}] \parbox{12.5cm}{Random event.}
	\item [\parbox{2cm}{$B_{1}$}] \parbox{12.5cm}{The first set in the partition $S$.}
	\item [\parbox{2cm}{$B_{2}$}] \parbox{12.5cm}{The second set in the partition $S$.}
	\item [\parbox{2cm}{$B_{i}$}] \parbox{12.5cm}{The $i$-th set in the partition $S$.}
	\item [\parbox{2cm}{$B_{j}$}] \parbox{12.5cm}{The $j$-th set in the partition $S$.}
	\item [\parbox{2cm}{$B_{K}$}] \parbox{12.5cm}{The last set in the partition $S$.}
	\item [\parbox{2cm}{$B(\alpha, \beta)$}] \parbox{12.5cm}{The normalising constant used in the \acs{PDF} of the Beta probability distribution.}
	\item [\parbox{2cm}{$B(\boldsymbol{\alpha})$}] \parbox{12.5cm}{The normalising constant used in the \acs{PDF} of the Dirichlet probability distribution.}
	\item [\parbox{2cm}{$B$}] \parbox{12.5cm}{Random event.}
	\item [\parbox{2cm}{$Ber$}] \parbox{12.5cm}{The Bernoulli probability distribution.}
	\item [\parbox{2cm}{$Beta$}] \parbox{12.5cm}{The Beta probability distribution.}
	\item [\parbox{2cm}{$Bin$}] \parbox{12.5cm}{The Binomial probability distribution.}
	\item [\parbox{2cm}{$c$}] \parbox{12.5cm}{Symmetric distribution constant.}
	\item [\parbox{2cm}{$Cat$}] \parbox{12.5cm}{The Categorical probability distribution.}
	\item [\parbox{2cm}{$Dir$}] \parbox{12.5cm}{The Dirichlet probability distribution.}
	\item [\parbox{2cm}{$E[\ln(x_{k})]$}] \parbox{12.5cm}{The expected value of the natural logarithm of the $k$-th dimension of the random variable.}
	\item [\parbox{2cm}{$E[\ln(x)]$}] \parbox{12.5cm}{The expected value of the natural logarithm of $x$.}
	\item [\parbox{2cm}{$E[x_{k}]$}] \parbox{12.5cm}{The expected value of the $k$-th dimension of the random variable.}
	\item [\parbox{2cm}{$E[x]$}] \parbox{12.5cm}{The expected value of the distribution of $x$.}
	\item [\parbox{2cm}{$f_{Ber}$}] \parbox{12.5cm}{The \acs{PMF} for the Bernoulli probability distribution.}
	\item [\parbox{2cm}{$f_{Beta}$}] \parbox{12.5cm}{The \acs{PDF} of the Beta probability distribution.}
	\item [\parbox{2cm}{$f_{Bin}$}] \parbox{12.5cm}{The \acs{PMF} for the Binomial probability distribution.}
	\item [\parbox{2cm}{$f_{Cat}$}] \parbox{12.5cm}{The \acs{PMF} for the Categorical probability distribution.}
	\item [\parbox{2cm}{$f_{Dir}$}] \parbox{12.5cm}{The \acs{PDF} of the Dirichlet probability distribution.}
	\item [\parbox{2cm}{$f_{Mul}$}] \parbox{12.5cm}{The \acs{PMF} for the Multinomial probability distribution.}
	\item [\parbox{2cm}{$f$}] \parbox{12.5cm}{General symbol used for functions. Used for likelihood functions, \acp{PMF} and \acp{PDF}.}
	\item [\parbox{2cm}{$F$}] \parbox{12.5cm}{The symbol used for a female outcome in the mice experiment.}
	\item [\parbox{2cm}{$g^{*}$}] \parbox{12.5cm}{Posterior density.}
	\item [\parbox{2cm}{$g$}] \parbox{12.5cm}{Prior density.}
	\item [\parbox{2cm}{$i$}] \parbox{12.5cm}{Index to track sets in partition $S$. Also used as a general index in random variables in $I$ dimensions.}
	\item [\parbox{2cm}{$I$}] \parbox{12.5cm}{The total number of independent, identical (iid) random events.}
	\item [\parbox{2cm}{$j$}] \parbox{12.5cm}{Index to track sets in partition $S$. Also used as a general index in random variables in $J$ dimensions.}
	\item [\parbox{2cm}{$k$}] \parbox{12.5cm}{Index used to denote class $k$ in $K$ classes. Also used as a general index in random variables in $K$ dimensions.}
	\item [\parbox{2cm}{$K$}] \parbox{12.5cm}{The total number of sets in the partition $S$. Also used to denote the number of classes in the Dirichlet probability distribution.}
	\item [\parbox{2cm}{$M[x_{k}]$}] \parbox{12.5cm}{The mode of the $k$-th dimension of the random variable.}
	\item [\parbox{2cm}{$M[x]$}] \parbox{12.5cm}{The mode of the distribution of $x$.}
	\item [\parbox{2cm}{$m$}] \parbox{12.5cm}{The marginal density function.}
	\item [\parbox{2cm}{$M$}] \parbox{12.5cm}{The symbol used for a male outcome in the mice experiment.}
	\item [\parbox{2cm}{$Mul$}] \parbox{12.5cm}{The Multinomial probability distribution.}
	\item [\parbox{2cm}{$N_{0}$}] \parbox{12.5cm}{Summary variable tracking the number of unsuccessful Boolean outcomes.}
	\item [\parbox{2cm}{$N_{1}$}] \parbox{12.5cm}{Summary variable tracking the number of successful Boolean outcomes.}
	\item [\parbox{2cm}{$N_{k}$}] \parbox{12.5cm}{Summary variable, denoting the number of times a category $k$ occurs over all trials in $N$.}
	\item [\parbox{2cm}{$N_{K}$}] \parbox{12.5cm}{Summary variable denoting the number of times category $k$ occurs.}
	\item [\parbox{2cm}{$n$}] \parbox{12.5cm}{Input parameter to the Gamma function.}
	\item [\parbox{2cm}{$N$}] \parbox{12.5cm}{The total number of events observed.}
	\item [\parbox{2cm}{$P$}] \parbox{12.5cm}{The probability of some event.}
	\item [\parbox{2cm}{$p$}] \parbox{12.5cm}{The statistical $p$-value.}
	\item [\parbox{2cm}{$S$}] \parbox{12.5cm}{The union of mutually exclusive subsets. Also referred to as the probability simplex.}
	\item [\parbox{2cm}{$x_{1}$}] \parbox{12.5cm}{The first dimension of the random variable. Also used as the first random event.}
	\item [\parbox{2cm}{$x_{i,k}$}] \parbox{12.5cm}{The $k$-th dimension of the $i$-th random event vector $\boldsymbol{x}$.}
	\item [\parbox{2cm}{$x_{i}$}] \parbox{12.5cm}{The $i$-th dimension of the random variable. Also used as the $i$-th random event.}
	\item [\parbox{2cm}{$x_{I}$}] \parbox{12.5cm}{The last dimension of the random variable with $I$ dimensions. Also used as the last random event in $I$ random events.}
	\item [\parbox{2cm}{$x_{j}$}] \parbox{12.5cm}{The $j$-th dimension of the random variable. Also used as the $j$-th random event.}
	\item [\parbox{2cm}{$x_{k}$}] \parbox{12.5cm}{The $k$-th dimension of the random variable. Also used as the $k$-th random event.}
	\item [\parbox{2cm}{$x_{K}$}] \parbox{12.5cm}{The last dimension of the random variable with $K$ dimensions. Also used as the last random event in $K$ random events.}
	\item [\parbox{2cm}{$x$}] \parbox{12.5cm}{Random variable used by various probability distributions. Also referred to as a realisation of an event that occurred as a results of a random process $X$.}
	\item [\parbox{2cm}{$Y_{1}$}] \parbox{12.5cm}{The first random event in $\boldsymbol{Y}$.}
	\item [\parbox{2cm}{$y_{1}$}] \parbox{12.5cm}{The realisation of the first random event in $\boldsymbol{Y}$.}
	\item [\parbox{2cm}{$y_{2}$}] \parbox{12.5cm}{The realisation of the second random event in $\boldsymbol{Y}$.}
	\item [\parbox{2cm}{$Y_{2}$}] \parbox{12.5cm}{The second random event in $\boldsymbol{Y}$.}
	\item [\parbox{2cm}{$Y_{i}$}] \parbox{12.5cm}{The $i$-th random event in $\boldsymbol{Y}$.}
	\item [\parbox{2cm}{$y_{i}$}] \parbox{12.5cm}{The realisation of the $i$-th random event in $\boldsymbol{Y}$.}
	\item [\parbox{2cm}{$Y_{N}$}] \parbox{12.5cm}{The last random event in $\boldsymbol{Y}$ with $N$ events.}
	\item [\parbox{2cm}{$y_{n}$}] \parbox{12.5cm}{The realisation of the $n$-th random event in $\boldsymbol{Y}$ with $N$ events.}
\end{description}

\section{Chapter~\ref{chap:bhh}: Bayesian Hyper-Heuristic}
\label{sec:symbols:bhh}

\begin{description}

	\item [\parbox{2cm}{$\alpha_{1}$}] \parbox{12.5cm}{The first dimension of the concentration parameter, $\boldsymbol{\alpha}$.}
	\item [\parbox{2cm}{$\alpha_{K}$}] \parbox{12.5cm}{The $k$-th dimension of the concentration parameter, $\boldsymbol{\alpha}$.}
	\item [\parbox{2cm}{$\alpha_{k}$}] \parbox{12.5cm}{The dimension of $\boldsymbol{\alpha}$ associated with class $k$.}
	\item [\parbox{2cm}{$\alpha$}] \parbox{12.5cm}{Concentration parameter for the Beta probability distribution.}
	\item [\parbox{2cm}{$\beta_{1}$}] \parbox{12.5cm}{The first dimension of the concentration parameter, $\boldsymbol{\beta}$.}
	\item [\parbox{2cm}{$\beta_{2}$}] \parbox{12.5cm}{The second dimension of the concentration parameter, $\boldsymbol{\beta}$.}
	\item [\parbox{2cm}{$\beta_{j,k}$}] \parbox{12.5cm}{The $j$-th entity's concentration parameter associated with heuristic $k$.}
	\item [\parbox{2cm}{$\beta$}] \parbox{12.5cm}{Concentration parameter for the Beta probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\alpha}$}] \parbox{12.5cm}{The concentration parameters for the heuristic probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\beta}$}] \parbox{12.5cm}{The concentration parameters for the entity-heuristic probability distributions.}
	\item [\parbox{2cm}{$\boldsymbol{\gamma}$}] \parbox{12.5cm}{The concentration parameters for the credit-heuristic probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\hat{\phi}}$}] \parbox{12.5cm}{The expected value of the entity-heuristic probabilities.}
	\item [\parbox{2cm}{$\boldsymbol{\hat{\psi}}$}] \parbox{12.5cm}{The expected value of the heuristic-credit-assignment probabilities.}
	\item [\parbox{2cm}{$\boldsymbol{\hat{\theta}}$}] \parbox{12.5cm}{The expected value of the heuristic selection probabilities.}
	\item [\parbox{2cm}{$\boldsymbol{\lambda}$}] \parbox{12.5cm}{Error vector.}
	\item [\parbox{2cm}{$\boldsymbol{\phi}$}] \parbox{12.5cm}{The entity-heuristic probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\psi}$}] \parbox{12.5cm}{The credit-heuristic probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\theta}$}] \parbox{12.5cm}{The heuristic probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{C}$}] \parbox{12.5cm}{The event of observing credit assignments upon meeting credit assignment criteria.}
	\item [\parbox{2cm}{$\boldsymbol{E}$}] \parbox{12.5cm}{The event of observing entities.}
	\item [\parbox{2cm}{$\boldsymbol{H}$}] \parbox{12.5cm}{The event of observing heuristics.}
	\item [\parbox{2cm}{$\epsilon$}] \parbox{12.5cm}{Error factor.}
	\item [\parbox{2cm}{$\eta$}] \parbox{12.5cm}{Learning rate.}
	\item [\parbox{2cm}{$\gamma_{0}$}] \parbox{12.5cm}{The zeroth dimension of the concentration parameter, $\boldsymbol{\gamma}$.}
	\item [\parbox{2cm}{$\gamma_{1,k}$}] \parbox{12.5cm}{The first dimension of the concentration parameter, $\boldsymbol{\gamma}$ associated with heuristic $k$.}
	\item [\parbox{2cm}{$\gamma_{1}$}] \parbox{12.5cm}{The first dimension of the concentration parameter, $\boldsymbol{\gamma}$.}
	\item [\parbox{2cm}{$\gamma_{2,k}$}] \parbox{12.5cm}{The second dimension of the concentration parameter, $\boldsymbol{\gamma}$ associated with heuristic $k$.}
	\item [\parbox{2cm}{$\gamma_{K}$}] \parbox{12.5cm}{The $k$-th dimension of the concentration parameter, $\boldsymbol{\gamma}$.}
	\item [\parbox{2cm}{$\hat{\phi}_{j,k}$}] \parbox{12.5cm}{The expected value of the probability of selecting heuristic $k$ for entity $j$.}
	\item [\parbox{2cm}{$\hat{\psi}_{k}$}] \parbox{12.5cm}{The expected value of the probability of selecting heuristic $k$ and observing credit assignments.}
	\item [\parbox{2cm}{$\hat{\theta}_{k}$}] \parbox{12.5cm}{The expected value of the probability of selecting heuristic $k$.}
	\item [\parbox{2cm}{$\lambda_{j}$}] \parbox{12.5cm}{The $j$-th dimension of the error vector, $\boldsymbol{\lambda}$.}
	\item [\parbox{2cm}{$\lambda_{J}$}] \parbox{12.5cm}{The last dimension of the error vector, $\boldsymbol{\lambda}$ in $J$ dimensions.}
	\item [\parbox{2cm}{$\mathbbm{1}_{0}$}] \parbox{12.5cm}{The indicator function yielding a failure/non-occurrence.}
	\item [\parbox{2cm}{$\mathbbm{1}_{1}$}] \parbox{12.5cm}{The indicator function yielding a success/occurrence.}
	\item [\parbox{2cm}{$\mathcal{A}(v)$}] \parbox{12.5cm}{Functional form of priors.}
	\item [\parbox{2cm}{$\mathcal{L}$}] \parbox{12.5cm}{The likelihood function.}
	\item [\parbox{2cm}{$\phi_{j,k}$}] \parbox{12.5cm}{The entity-heuristic selection probability for entity $j$ and heuristic $k$.}
	\item [\parbox{2cm}{$\psi_{k}$}] \parbox{12.5cm}{The successful credit assignment probability for heuristic $k$.}
	\item [\parbox{2cm}{$\theta_{k}$}] \parbox{12.5cm}{The heuristic selection probability for heuristic $k$.}
	\item [\parbox{2cm}{$A$}] \parbox{12.5cm}{A random event.}
	\item [\parbox{2cm}{$B$}] \parbox{12.5cm}{A random event.}
	\item [\parbox{2cm}{$Beta$}] \parbox{12.5cm}{The Beta probability distribution.}
	\item [\parbox{2cm}{$Bin$}] \parbox{12.5cm}{The Binomial probability distribution.}
	\item [\parbox{2cm}{$c_{1}$}] \parbox{12.5cm}{Successful credit allocation.}
	\item [\parbox{2cm}{$c_{i,k}$}] \parbox{12.5cm}{The realisation of the $i$-th event of observing a credit assignment $i$ to heuristic $k$.}
	\item [\parbox{2cm}{$c_{i}$}] \parbox{12.5cm}{The $i$-th event of observing credit assignments.}
	\item [\parbox{2cm}{$Dir$}] \parbox{12.5cm}{The Dirichlet probability distribution.}
	\item [\parbox{2cm}{$e_{i,j,k}$}] \parbox{12.5cm}{The realisation of the $i$-th event of observing heuristic $k$ for selected for entity $j$.}
	\item [\parbox{2cm}{$e_{i}$}] \parbox{12.5cm}{The $i$-th event of observing entities.}
	\item [\parbox{2cm}{$e_{j}$}] \parbox{12.5cm}{The $j$-th entity.}
	\item [\parbox{2cm}{$E[\phi_{j,k}]$}] \parbox{12.5cm}{The expected value of the probability of selecting heuristic $k$ for entity $j$.}
	\item [\parbox{2cm}{$E[\psi_{k}]$}] \parbox{12.5cm}{The expected value of the probability of selecting heuristic $k$ and observing credit assignments.}
	\item [\parbox{2cm}{$E[\theta_{k}]$}] \parbox{12.5cm}{The expected value of the probability of selecting heuristic $k$.}
	\item [\parbox{2cm}{$h_{i,k}$}] \parbox{12.5cm}{The realisation of the $i$-th event of observing heuristic $k$.}
	\item [\parbox{2cm}{$h_{i}$}] \parbox{12.5cm}{The $i$-th event of observing heuristics.}
	\item [\parbox{2cm}{$h_{k}$}] \parbox{12.5cm}{The $k$-th heuristic.}
	\item [\parbox{2cm}{$i$}] \parbox{12.5cm}{Index generally associated with event/run $i$.}
	\item [\parbox{2cm}{$I$}] \parbox{12.5cm}{The maximum number of instances in the replay window.}
	\item [\parbox{2cm}{$j$}] \parbox{12.5cm}{Index generally associated to entity $j$.}
	\item [\parbox{2cm}{$J$}] \parbox{12.5cm}{The entity pool (population) size.}
	\item [\parbox{2cm}{$k$}] \parbox{12.5cm}{Index generally associated to heuristic $k$.}
	\item [\parbox{2cm}{$K$}] \parbox{12.5cm}{The heuristic pool size.}
	\item [\parbox{2cm}{$L$}] \parbox{12.5cm}{The number of credit assignment output classes.}
	\item [\parbox{2cm}{$LSE$}] \parbox{12.5cm}{The log-sum-exp trick/function.}
	\item [\parbox{2cm}{$Mult$}] \parbox{12.5cm}{The Multinomial probability distribution.}
	\item [\parbox{2cm}{$N_{0,k}$}] \parbox{12.5cm}{The count of occurrences of the events $c_{i}$ taking on a failure.}
	\item [\parbox{2cm}{$N_{1,k}$}] \parbox{12.5cm}{The count of occurrences of the events $c_{i}$ taking on a success.}
	\item [\parbox{2cm}{$N_{j,k}$}] \parbox{12.5cm}{A summary variable denoting the count of occurrences of the events $e_{i}$ taking on class $j$ and $h_{i}$ taking on class $k$ in $I$ independent, identical runs.}
	\item [\parbox{2cm}{$N_{j}$}] \parbox{12.5cm}{The count of the occurrences of observing entity $j$.}
	\item [\parbox{2cm}{$N_{k}$}] \parbox{12.5cm}{A summary variable denoting the count of occurrences of the event $h_{i}$ taking on class $k$ in $I$ independent, identical runs.}
	\item [\parbox{2cm}{$N$}] \parbox{12.5cm}{The maximum number of random events observed.}
	\item [\parbox{2cm}{$S$}] \parbox{12.5cm}{The probability simplex.}
	\item [\parbox{2cm}{$t$}] \parbox{12.5cm}{Time step.}
	\item [\parbox{2cm}{$\lambda_{1}$}] \parbox{12.5cm}{The first dimension of the error vector, $\boldsymbol{\lambda}$.}
\end{description}

\section{Chapter~\ref{chap:methodology}: Methodology}
\label{sec:symbols:methodology}

\begin{description}
	\item [\parbox{2cm}{$\alpha$}] \parbox{12.5cm}{Used as a scaling parameter for the \acs{LReLU} activation function. Also used as a threshold parameter for the statistical significance tests.}
	\item [\parbox{2cm}{$\eta$}] \parbox{12.5cm}{Learning rate.}
	\item [\parbox{2cm}{$\mathbb{R}$}] \parbox{12.5cm}{Real-number space.}
	\item [\parbox{2cm}{$v_{ij}$}] \parbox{12.5cm}{The $j$-th dimension of the velocity vector of the $i$-th particle in the population implemented by \acs{PSO}.}
	\item [\parbox{2cm}{$x_{ij}$}] \parbox{12.5cm}{The $j$-th dimension of the position vector of the $i$-th particle in the population implemented by \acs{PSO}.}
\end{description}

\section{Chapter~\ref{chap:results}: Results}
\label{sec:symbols:results}

\begin{description}
	\item [\parbox{2cm}{$\alpha_{0}$}] \parbox{12.5cm}{The zeroth dimension of the concentration parameter $\boldsymbol{\alpha}$ and is associated to heuristic $0$ (\acs{SGD}).}
	\item [\parbox{2cm}{$\alpha_{6}$}] \parbox{12.5cm}{The sixth dimension of the concentration parameter $\boldsymbol{\alpha}$ and is associated to heuristic $6$ (\acs{Adam}).}
	\item [\parbox{2cm}{$\alpha_{7}$}] \parbox{12.5cm}{The seventh dimension of the concentration parameter $\boldsymbol{\alpha}$ and is associated to heuristic $7$ (\acs{PSO}).}
	\item [\parbox{2cm}{$\alpha_{8}$}] \parbox{12.5cm}{The eight dimension of the concentration parameter $\boldsymbol{\alpha}$ and is associated to heuristic $8$ (\acs{GA}).}
	\item [\parbox{2cm}{$\alpha_{i}$}] \parbox{12.5cm}{The $i$-th dimension of the concentration parameter $\boldsymbol{\alpha}$ and is associated to heuristic $i$.}
	\item [\parbox{2cm}{$\boldsymbol{\alpha}$}] \parbox{12.5cm}{The concentration parameters for the heuristic probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\phi}$}] \parbox{12.5cm}{The entity-heuristic probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\psi}$}] \parbox{12.5cm}{The credit-heuristic probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{\theta}$}] \parbox{12.5cm}{The heuristic probability distribution.}
	\item [\parbox{2cm}{$\boldsymbol{C}$}] \parbox{12.5cm}{The event of observing credit assignments upon meeting credit assignment criteria.}
	\item [\parbox{2cm}{$\boldsymbol{E}$}] \parbox{12.5cm}{The event of observing entities.}
	\item [\parbox{2cm}{$\boldsymbol{H}$}] \parbox{12.5cm}{The event of observing heuristics.}
	\item [\parbox{2cm}{$\theta_{0}$}] \parbox{12.5cm}{The zeroth dimension of the heuristic selection probabilities $\boldsymbol{\theta}$ and is associated to heuristic $0$ (\acs{SGD}).}
	\item [\parbox{2cm}{$\theta_{6}$}] \parbox{12.5cm}{The sixth dimension of the heuristic selection probabilities $\boldsymbol{\theta}$ and is associated to heuristic $6$ (\acs{Adam}).}
	\item [\parbox{2cm}{$\theta_{7}$}] \parbox{12.5cm}{The seventh dimension of the heuristic selection probabilities $\boldsymbol{\theta}$ and is associated to heuristic $7$ (\acs{PSO}).}
	\item [\parbox{2cm}{$\theta_{8}$}] \parbox{12.5cm}{The eight dimension of the heuristic selection probabilities $\boldsymbol{\theta}$ and is associated to heuristic $8$ (\acs{GA}).}
	\item [\parbox{2cm}{$c_{1}$}] \parbox{12.5cm}{Successful credit allocation.}
	\item [\parbox{2cm}{$Cat$}] \parbox{12.5cm}{The Categorical probability distribution.}
	\item [\parbox{2cm}{$Dir$}] \parbox{12.5cm}{The Dirichlet probability distribution.}
	\item [\parbox{2cm}{$e_{0}$}] \parbox{12.5cm}{Entity $0$.}
	\item [\parbox{2cm}{$h_{0}$}] \parbox{12.5cm}{Heuristic $0$ (\acs{SGD}).}
	\item [\parbox{2cm}{$h_{6}$}] \parbox{12.5cm}{Heuristic $6$ (\acs{Adam}).}
	\item [\parbox{2cm}{$h_{7}$}] \parbox{12.5cm}{Heuristic $7$ (\acs{PSO}).}
	\item [\parbox{2cm}{$h_{8}$}] \parbox{12.5cm}{Heuristic $8$ (\acs{GA}).}
	\item [\parbox{2cm}{$i$}] \parbox{12.5cm}{General index associated with event/run number.}
	\item [\parbox{2cm}{$K$}] \parbox{12.5cm}{Heuristic pool size.}
\end{description}