### 2025-01-13 dissertation/preamble/supervisors.tex

\pagestyle{empty}
\vspace*{\fill}
\noindent
\parbox{4cm}{\textbf{Supervisor:}}\parbox{12cm}{Dr. A.S. Bosman}\\*[2mm]
\parbox{4cm}{\textbf{~}}\parbox{12cm}{University of the Pretoria}\\*[2mm]
\parbox{4cm}{\textbf{~}}\parbox{12cm}{Department of Computer Science}\\*[1cm]
\parbox{4cm}{\textbf{Supervisor:}}\parbox{12cm}{Prof. C.W. Cleghorn}\\*[2mm]
\parbox{4cm}{\textbf{~}}\parbox{12cm}{University of the Witwatersrand}\\*[2mm]
\parbox{4cm}{\textbf{~}}\parbox{12cm}{Department of Computer Science}\\*[1cm]
\parbox{4cm}{\textbf{Supervisor:}}\parbox{12cm}{Prof. A.P. Engelbrecht}\\*[2mm]
\parbox{4cm}{\textbf{~}}\parbox{12cm}{Stellenbosch University}\\*[2mm]
\parbox{4cm}{\textbf{~}}\parbox{12cm}{Department of Industrial Engineering, Computer Science Division}\\*[1cm]
\parbox{4cm}{\textbf{Degree:}}\parbox{12cm}{Master of Science, Computer Science (Artificial Intelligence)}
\newpage